### 2019-03-31 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Version: to PRApplied

\documentclass[aps,pra,amsmath,amssymb,onecolumn,superscriptaddress,showpacs,floatfix,]{revtex4-1}

\bibliographystyle{apsrev4-1}

\usepackage{ifpdf}
\newif\ifpdf
\ifx\pdfoutput\undefined
\pdffalse
\else
\pdfoutput=1
\pdftrue
\fi
\ifpdf
\usepackage{graphicx}
\usepackage{epstopdf}
% \DeclareGraphicsRule{.eps}{pdf}{.pdf}{`epstopdf #1}
% \DeclareGraphicsRule{.png}{png}{.pdf}{`epstopdf #1}
\pdfcompresslevel=9
\pdfobjcompresslevel=9
\pdfminorversion=9
\else
\usepackage{graphicx}
% \DeclareGraphicsRule{.jpg}{jpg}{}{}
\fi

\usepackage{bm}
\usepackage{epsfig}
\usepackage[usenames]{color}
\usepackage{soul}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{amsmath}
\newcommand{\Cs}{\mathit C_{\scriptscriptstyle{\Sigma}}}
\newcommand{\Rs}{\mathit R_{\scriptscriptstyle{\Sigma}}}
\DeclareMathOperator{\sign}{sign}

%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\begin{document}

%\titlefigure{Fig1_fin.pdf}

\title{Response time of a plasmonic distributed feedback laser in large-signal modulation regime}

\author{N.~E.~Nefedkin}
\email{nefedkin@phystech.edu}
\affiliation{Dukhov Research Institute of Automatics (VNIIA), 22 Sushchevskaya, Moscow 127055, Russia}
\affiliation{Moscow Institute of Physics and Technology, Moscow 141700, Russia}
\affiliation{Institute for Theoretical and Applied Electromagnetics, 13 Izhorskaya, Moscow 125412, Russia}

\author{A.~A.~Zyablovsky}
\email{zyablovskiy@mail.ru}
\affiliation{Dukhov Research Institute of Automatics (VNIIA), 22 Sushchevskaya, Moscow 127055, Russia}
\affiliation{Moscow Institute of Physics and Technology, Moscow 141700, Russia}

\author{E.~S.~Andrianov}
\affiliation{Dukhov Research Institute of Automatics (VNIIA), 22 Sushchevskaya, Moscow 127055, Russia}
\affiliation{Moscow Institute of Physics and Technology, Moscow 141700, Russia}

\author{A.~A.~Pukhov}
\affiliation{Dukhov Research Institute of Automatics (VNIIA), 22 Sushchevskaya, Moscow 127055, Russia}
\affiliation{Moscow Institute of Physics and Technology, Moscow 141700, Russia}
\affiliation{Institute for Theoretical and Applied Electromagnetics, 13 Izhorskaya, Moscow 125412, Russia}

\author{A.~P.~Vinogradov}
\affiliation{Dukhov Research Institute of Automatics (VNIIA), 22 Sushchevskaya, Moscow 127055, Russia}
\affiliation{Moscow Institute of Physics and Technology, Moscow 141700, Russia}
\affiliation{Institute for Theoretical and Applied Electromagnetics, 13 Izhorskaya, Moscow 125412, Russia}

\date{\today}% It is always \today, today,
%  but any date may be explicitly specified

\begin{abstract}

	The time response to an external signal is the main characteristic of optoelectronic devices which determines their maximum modulation frequency.
	The use of plasmonic structures gives the opportunity to significantly reduce the response time.
	In this paper, we study the temporal dynamics of a plasmonic distributed feedback laser consisting of a two-dimensional metallic structure coated with a layer of the active medium.
	Combining numerical simulations with an analytical evaluation we show that the response time in the large-signal modulation regime strongly depends on the size of the pump beam.
	We demonstrate that the interaction of a large number of modes of the plasmonic structure with the pumped active medium leads to the formation of a single collective mode.
	If the size of the pump beam is larger than the propagation length of electromagnetic waves in the plasmonic structure, the collective mode is localized in the pumped area.
	In this case, the response time slightly increases with the size of the pump beam.
	In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
	We demonstrate that there is an optimal size of the pump beam for which the modulation speed is a maximum, and for typical experimental parameters can achieve $1$ THz.
	The obtained result opens the way to increasing both the modulation speed and the energy efficiency of plasmonic laser devices.

\end{abstract}

\maketitle

\section*{Introduction}
\label{sec:intro}
For many optoelectronic applications, such as the modulation of optical signals and sensing, lasers with an ultrafast modulation response are required \cite{daskalakis2018nanolett, ZhouNatNano}.
Such devices often operate in a large-signal modulation scheme where a pumping pulse switches the laser between the subthreshold regime and the above-threshold regime.
In such a scheme the response of the laser to a pumping pulse can be divided into three stages \cite{altug2006ultrafast, altug2008lpr, SiegmanLasers}.
First, the response is initiated by spontaneous emission in the active medium, which creates the electromagnetic (EM) field in the cavity.
In the second stage, the stimulated emission begins to prevail over the spontaneous emission and the intensity of the EM field in the cavity rapidly increases, while the population inversion of the active medium decreases.
When the population inversion has vanished, the growth of the EM field intensity ceases and the third stage begins.
The EM field intensity drops due to losses in the cavity.

During the first stage, the intensity of the EM field in the cavity is negligible, and only in the second stage does a significant response to the external pumping pulse appear.
Hence the duration of the first stage determines the delay time of the output pulse.
It can be reduced by accelerating the spontaneous emission by an increase of the Purcell factor \cite{altug2006ultrafast, altug2008lpr}.
To achieve a high Purcell factor one can use a large Rabi constant of the interaction between the electric field and the active medium.
In turn, the increase of the Rabi constant leads to the rise of the rate of stimulated emission and a decrease of the duration of the second stage.
The duration of the third stage is determined by the relaxation rate of the EM field in the cavity.
Thus, to create a short output pulse one has to work in the high-Purcell regime in order to reduce the delay time, and use cavities with a high relaxation rate to shorten the third stage and narrow the width of the pulse.

The maximal modulation frequency of surface emitting lasers based on dielectric structures is limited to several tens of GHz \cite{coldren2012diode, scott1994high}.
This frequency can be significantly increased if one use metallic plasmonic structures.
The localization of the EM field at the nanoscale results in a strong interaction with the atoms of the active medium and, as a consequence, a high Purcell factor.
Losses in the metal cause a high relaxation rate of the EM field of the plasmonic structure.
Both of these factors lead to an increase of the modulation speed.
At the same time, the use of periodic structures enables achieving desirable radiation losses and beam directionality of plasmonic lasers.
These arguments have stimulated significant interest in the creation and study of lasers based on plasmonic nanoparticle arrays or perforated films \cite{BeijnumPRL,SchokkerPRB,ZhouNatNano,MengLPR}, particularly in the large-signal modulation regime \cite{daskalakis2018nanolett}.

Periodicity inevitably results in the multimode nature of these lasers \cite{BeijnumPRL,TennerJOpt}.
As a result, the excited active medium emits in many modes simultaneously.
This leads to a decrease in the energy in each mode and a delay of the second stage in which the stimulated emission dominates.
Consequently, the total response time increases, which may limit the advantages of the use of plasmonic structures.
Thus, the question arises about the possibility of getting over the difficulties associated with the multimode nature of plasmonic lasers.

In our paper, we study the dynamics in the large-signal modulation regime of a plasmonic distributed feedback (DFB) laser that consists of a gold film perforated by a periodic array of holes, and is coated with a layer of an active medium.
We examine the behavior of the laser for different sizes of the pump beam.
Combining numerical simulations of the laser dynamics within the framework developed in Ref. \cite{Zyablovsky2017approach,nefedkin2018acsphot} with an analytical evaluation, we show that the response time strongly depends on the size of the pump beam.
This dependence is due to the complex structure of the modes of the DFB laser.
The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump.
If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
The response time in this case reaches picoseconds and slightly increases with the size of the pump beam.
In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
We demonstrate that there is an optimal size of the pump beam, for which the response time reaches a minimum.
For typical experimental parameters, the optimal size of the pump beam is $\sim 15 \mu$m and the minimal response time is $\sim 1$ $\text{ps}$.
The obtained value of the response time is in agreement with a recent experiment \cite{daskalakis2018nanolett} where it has been shown that plasmonic lasers have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime.
Thus, the optimization of the pump beam size makes it possible to increase the modulation speed as well as the energy efficiency of plasmonic DFB laser devices.

\begin{figure}[!h]
	\centering
	\begin{subfigure}[h]{0.6\linewidth}
		\includegraphics[width=\linewidth]{Fig1.png}
		%\caption{Coffee.}
	\end{subfigure}
	\caption{The sketch of the two-dimensional plasmonic distributed feedback laser. The diameter of the holes is $160$ $\text{nm}$, the distance between holes is $470$ $\text{nm}$, the thickness of the Au film is $100$ $\text{nm}$.}
	\label{fig1}
\end{figure}

\section*{Temporal dynamics of a two-dimensional plasmonic DFB laser}
A conventional plasmonic DFB laser consists of a periodic plasmonic structure, e.g., a metallic film perforated by an array of holes \cite{BeijnumPRL,DorofeenkoOptExp,MengLPR,nefedkin2018acsphot,TennerJOpt,TennerACSPhot,Zyablovsky2017Optimum,Melentiev2017nanolaser} or two-dimensional arrays of plasmonic nanoparticles \cite{ZhouNatNano,SchokkerPRB,daskalakis2018nanolett,SchokkerACSPhot,YangNatComm,YangACSNano,hakala2017lasing,wang2017band}, and an active medium.
The eigenmodes of such a system are Bloch modes which are distributed over the whole surface of the laser.
To overcome losses in the metal, an external pumping (optical or electrical) of the active medium is used.
In the case of optical pumping, the diameter of the pump beam is usually smaller than the size of the periodic plasmonic structure \cite{TennerACSPhot,ZhouNatNano,daskalakis2018nanolett}, see Fig.~\ref{fig1}a.
In such a case, the EM field is excited inside the pumped area and is formed by a large number of Bloch modes \cite{TennerACSPhot,nefedkin2018acsphot}.
To describe the multimode dynamics of the plasmonic DFB laser we use an approach which has been proposed in Ref. \cite{Zyablovsky2017approach} for considering ultrafast phenomena in dispersive dissipative media.
This approach is based on the following equations:
\begin{equation}\label{eq1}
\frac{{d{n_{jk}}}}{{dt}} =  - \left( {{\gamma _j} + {\gamma _k}} \right){n_{jk}} + i\left( {{\omega _j} - {\omega _k}} \right){n_{jk}} + \sum\limits_{m=1}^{N_{\text{atom}}} {\left( {{\Omega _{km}}{\varphi _{jm}} + \Omega _{jm}^*\varphi _{km}^*} \right)} 
\end{equation}
\begin{equation}\label{eq2}
\frac{{d{D_m}}}{{dt}} =  - {\gamma _D}\left( {1 + {D_m}} \right)  - 2\sum\limits_{j=1}^{N_{\text{mode}}} {\left( {{\Omega _{jm}}{\varphi _{jm}} + \Omega _{jm}^*\varphi _{jm}^*} \right)} 
\end{equation}
\begin{equation}\label{eq3}
\frac{{d{\varphi _{jm}}}}{{dt}} =  -\gamma_{jm}^{\varphi} {\varphi _{jm}} + i\left( {{\omega _j^a} - {\omega _\sigma }} \right){\varphi _{jm}} + \frac{{\Omega _{jm}^*}}{2}\left( {{D_m} + 1} \right) + \sum\limits_{l=1}^{N_{\text{mode}}} {\Omega _{lm}^*{n_{jl}}{D_m}} , 
\end{equation}
Here, $n_{jj}$ is the photon number in the $j$th Bloch mode, $n_{jk}$ is an interference term responsible for the flow of photons from the $k$th to the $j$th mode, when $j \neq k$.
$D_m$ is the population inversion of the $m$th atom and $\varphi _{jm} = - i \langle \hat{a} _j^{+} \hat{\sigma} _m \rangle$ describes the energy flow between the $m$th atom and $j$th mode.
The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the eigenfrequencies of the modes; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms, $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
$\Omega _{jm}$ is the Rabi constant of coupling between the $j$th mode and the $m$th atom.
These equations generalize the rate equations widely used in laser theory \cite{SiegmanLasers}.
In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the atoms, and the interference between the EM fields of the different modes, $n_{jk}$.

For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes, see Fig.~\ref{fig1}.
Such a structure has been realized in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
We consider the response of the system to ultrashort pulse pumping of the active medium.
Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}).

Using Eqs.~(\ref{eq1})--(\ref{eq3}), we simulate the time response to an ultrashort pump pulse for different diameters of the pump beam.
In Fig.~\ref{fig2}a the dependence of the photon number on time is shown.
As mentioned in the Introduction, there are three stages of the response of the system.
In the first stage, the response of the system is determined by the spontaneous emission in the active medium.
The photon number in the plasmonic DFB laser is small.
For the system under consideration, the duration of this stage, i.e., the delay time, turns out to be less than a picosecond.
In the second stage, the photon number in the plasmonic laser rapidly increases due to stimulated emission in the active medium.
This growth ceases with the saturation of the population inversion in the active medium.
In the third stage, the photon number in the system decreases exponentially with time.
We identify the total response time with the time when the number of photons in the system has decreased by a factor of ten from its maximum value.
This time is about one picosecond.

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.45\linewidth}
		\includegraphics[width=\linewidth]{Fig2a.eps}
		%\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.45\linewidth}
		\includegraphics[width=\linewidth]{Fig2b.eps}
		%\caption{}
	\end{subfigure}
	\caption{(a) The dependence of the total photon number on the time at the different diameters of the pump beam. The diameters are equal to 20 $\mu$m (black line), 10 $\mu$m (blue line), and 5 $\mu$m (red line). (b) The dependence of the modulation times on the diameters of the pump beam. The blue line depicts the delay time; the green line depicts the time of the end of the second stage; the red line depicts the total response time.}
	\label{fig2}
\end{figure}

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3a.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3b.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3c.png}
%		\caption{}
	\end{subfigure}
	\\
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3d.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3e.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3f.png}
%		\caption{}
	\end{subfigure}
	\caption{Field distribution of the collective mode when area of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a) -(c) and $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f). The time moments in (a)-(c) and (d)-(f) correspond to circles on the red curve and the black curve in Fig.\ref{fig2}a, respectively.}
	\label{fig3}	
\end{figure}

The dependence of the total photon number on time is qualitatively the same for different diameters of the pump beam.
However, the response times manifest a non-monotonic dependence on the size of the pump beam, see Fig.~\ref{fig2}b.
It can be seen that there is an optimal size of the pump beam at which the response time is a minimum.
For the considered plasmonic DFB laser the optimal diameter of the pump beam is about 15 $\mu$m.
As the pump beam diameter decreases, the response time increases rapidly.
With an increase in the diameter of the pump beam, the response time changes slowly.
This kind of behavior of the plasmonic DFB laser takes place at different initial values of the population inversion of the active atoms.
Thus, there is an optimal diameter of the pump beam, $d_{\text{beam}}^*$, at which an ultrafast light modulation in the plasmonic DFB laser is achieved.

\section*{Single-mode approximation}

The existence of an optimal size of the pump beam can be explained with the example of a single-mode model.
To justify the single-mode approximation, we calculate the distribution of the total electric field over the area of the DFB laser at different moments of time, see Fig.~\ref{fig3}.
When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
In the opposite case, $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
However, in both cases, the spatial distribution of the total electric field remains almost unchanged during its temporal evolution.
The time independence of the EM field distribution allows us to consider the plasmonic laser as a single mode laser.

Based on this observation, we assume that the part of the active medium pumped by an external beam effectively excites a single collective mode, which is formed by the Bloch modes of the plasmonic structure.
In this case, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
\begin{gather} 
\dot n =  - 2{\gamma _a}n + 2{N_{\text{at}}}{\Omega _R}\varphi \label{ZE1}
\\
\dot \varphi = - \left( {{\gamma _a} + {\gamma _\sigma }} \right)\varphi + {\Omega _R}\left( {2nD + D + 1} \right)/2 \label{ZE2}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4{\Omega _R}\varphi, \label{ZE3}
\end{gather}
where $n$ is the number of photons in the collective mode, $D$ is the mean population inversion of pumped atoms of the active medium, $\varphi$ describes the mean energy flow from the active medium to the collective mode, and $N_{\text{at}}$ is the number of atoms of the active medium lying in the collective mode.
We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.
Thus, the phase relaxation rate is much larger than other relaxation rates, $\gamma_\sigma \gg \gamma_a, \gamma_D$.
For this reason, one can adiabatically eliminate $\varphi$ from Eqs.~(\ref{ZE1})--(\ref{ZE3}) and obtain the conventional rate equations
\begin{gather} 
\dot n =  - {\gamma _a}n + 2{N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
\end{gather}
As initial condition we take $D(0) = 1$ and $n(0) = 0$, i.e., all atoms are inverted and there are no photons in the mode.
The temporal dynamics of such a system is well-known \cite{SiegmanLasers}; it is shown in Fig.~\ref{fig4}.

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.45\linewidth}
		\includegraphics[width=\linewidth]{Fig4a.eps}
		%\caption{}
	\end{subfigure}
	\begin{subfigure}[h!]{0.45\linewidth}
		\includegraphics[width=\linewidth]{Fig4b.eps}
		%\caption{}
	\end{subfigure}
	\caption{The dependences of the photon number (a) and the population inversion of atom (b) on time obtained from the single-mode model, Eqs. (\ref{RE1})-(\ref{RE2}); the vertical dashed lines denote the delay time, $t_{\text{delay}}$, obtained with Eq. (\ref{tdel}).
	}
	\label{fig4}
\end{figure}

The dynamics of the photon number in the collective mode obtained with Eqs.~(\ref{RE1})--(\ref{RE2}) is similar to the behavior of the total photon number in the real plasmonic DFB laser, cf.  Fig.~\ref{fig4}a and Fig.~\ref{fig2}a.
We observe three stages of the laser response, which are quantitatively similar to the stages in the real plasmonic DFB laser.

In the first stage, $D$ is constant and $n$ reaches stationary states at fixed $D$.
Under this assumption the solution of Eq.~(\ref{RE1}) is
\begin{equation} 
n\left( t \right) = \frac{{2{N_{\text{at}}}\Omega _R^2\left( {D + 1} \right)/{\gamma _\sigma }}}{{4{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma } - {\gamma _a}}}\left( {\exp \left( {4{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma } - {\gamma _a}} \right)t - 1} \right) = \frac{{\xi \left( {D + 1} \right)}}{{2\left( {\xi-1} \right)}}\left( {\exp \left( {\xi D - 1} \right){\gamma _a}t - 1} \right), \label{ph}
\end{equation}
where $\xi={4{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma \gamma_a }}$ is the inverse threshold value of the pumping in the CW regime \cite{ScullyQO}.
At the initial moment, $D \simeq 1$ and we have  
\begin{equation} 
n\left( t \right) = \frac{\xi }{{\left( {\xi - 1} \right)}}\left( {\exp \left( {\xi  - 1} \right){\gamma _a}t - 1} \right). \label{ph0}
\end{equation}
Substituting Eq.~(\ref{ph0}) into Eq.~(\ref{RE2}) one obtains
\begin{equation} 
\dot D =  - {\gamma _D}\left( {1 + \eta } \right)\left( {D + 1} \right) - {\gamma _D}\frac{{2\xi \eta D}}{{\left( {\xi  - 1} \right)}}\left( {\exp \left( {\xi  - 1} \right){\gamma _a}t - 1} \right), \label{inv0}
\end{equation}
where we have put $\eta = 4 \Omega_R^2 /\gamma_D \gamma_\sigma$.

The second stage corresponds to the time when the exponential term in Eq.~(\ref{inv0}) is dominant, and the dynamics is strongly nonlinear.
This nonlinear term describes the stimulated energy flow from the atoms to the collective mode.
In this case, from Eq.~(\ref{inv0}) we obtain the equation
\begin{equation} 
\dot D =  - {\gamma _D}\frac{{2\xi \eta D}}{{\left( {\xi  - 1} \right)}}{\exp \left( {\xi  - 1} \right){\gamma _a}t} \label{invnl1}
\end{equation}
which has the solution
\begin{equation} 
D\left( t \right) \simeq \exp \left( { - \frac{{2\xi \eta {\gamma _D}}}{{{{\left( {\xi  - 1} \right)}^2}{\gamma _a}}}\exp \left( {\xi  - 1} \right){\gamma _a}t} \right) = \exp \left( { - \frac{{2{\xi ^2}}}{{{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}}}\exp \left( {\xi  - 1} \right){\gamma _a}t} \right).
\end{equation}
The characteristic time for $D$ to change is 
\begin{equation} 
{t_{delay}} = \gamma _a^{ - 1}{\left( {\xi  - 1} \right)^{ - 1}}\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/2{\xi ^2}} \right). \label{tdele}
\end{equation}
At this time, a sharp decrease of $D$ appears, see Fig.~\ref{fig4}b.
Note that the assumption that the evolution of $D$ is much slower than the evolution of $n$ is valid when $t_{delay} \gg \gamma_a^{-1} \xi^{-1}$, i.e., $\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/2{\xi ^2}} \right) \gg 1$.
When $\xi \gg 1$, the assumption is correct if $N_{\text{at}} \gg 1$, which is true for the considered system.

For $\xi \gg 1$, Eq.~(\ref{tdele}) simplifies to  
\begin{equation} 
{t_{\text{delay}}} = {\gamma _\sigma }\ln \left( {{N_{\text{at}}}/2} \right)/4{N_{\text{at}}}\Omega _R^2. \label{tdel}
\end{equation}

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.45\linewidth}
		\includegraphics[width=\linewidth]{Fig5a.eps}
		%\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.45\linewidth}
		\includegraphics[width=\linewidth]{Fig5b.eps}
		%\caption{}
	\end{subfigure}
	\caption{(a) The dependence of the diameter of the collective mode on the pump beam diameter (blue line). The black dashed line denotes the pump beam diameter. (b) The dependence of the delay time on the pump beam diameter. The blue solid line is obtained from numerical simulation of Eqs. (\ref{eq1})--(\ref{eq3}), the orange thick line is obtained from Eq. (\ref{tdel}).
	}
	\label{fig5}
\end{figure}

Note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
The superradiance problem and the problem that we investigate here have the same physical grounds.
Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).

The figure \ref{fig5}b shows that the analytical approximation, Eq.~(\ref{tdel}), agrees well with the numerical simulation of the exact model, Eqs. (\ref{eq1})--(\ref{eq3}).
Let us analyze the dependence of the delay time, Eq.~(\ref{tdel}), on the size of the pumped area.
The number of atoms depends quadratically on the diameter $d_{\text{beam}}$ of the pump beam, i.e., $N_{\text{at}} \sim d_{\text{beam}}^2$.
In turn, the square of the Rabi constant is inversely proportional to the square of the diameter $d_{\text{mode}}^2$ of the collective mode, i.e., $\Omega_{R}^{2} \sim 1/d_{\text{mode}}^2$.
So, we have $t_{\text{delay}} \sim d_{\text{mode}}^2\ln \left( d_{\text{beam}}^2 \right ) /d_{\text{beam}}^2$.
The dependence of $d_{\text{mode}}$ on $d_{\text{beam}}$ is shown in Fig.~\ref{fig5}a.

It can be seen that there is a critical value, $d_{\text{beam}}^*$, of the diameter of the pump beam, which divides two different cases.
When $d_{\text{beam}}>d_{\text{beam}}^*$, the electric field of the collective mode is localized inside the pump beam, see Figs.~\ref{fig3}(d)--(f).
As a consequence, the mode diameter is proportional to the beam diameter, i.e., $d_{\text{mode}} \sim d_{\text{beam}}$, cf. Fig.~\ref{fig5}a.
For this case, from Eq.~(\ref{tdel}) we have $t_{\text{delay}} \sim \ln d_{\text{beam}}$.
This dependence is in agreement with a slow growth of the delay time of the DFB laser, see Fig\ref{fig2}b.
When $d_{\text{beam}}<d_{\text{beam}}^*$, the electric field of the collective mode is localized outside the pump beam, see Figs.~\ref{fig3}(a)--(c), and the mode size is determined by the characteristic propagation length of the EM field in the unpumped region.
As a result, the mode diameter does not depend on the beam diameter, and $d_{\text{mode}} \sim \text{const}$, see Fig.~\ref{fig5}a.
In this case, $t_{\text{delay}} \sim \ln(d_{\text{beam}}^2)/d_{\text{beam}}^2$.
This is in agreement with an increase of the delay time  when the pump beam diameter of the plasmonic DFB laser is reduced, see Fig.~\ref{fig2}b.
Note that the critical value $d_{\text{beam}}^*$ corresponds with the characteristic propagation length of the EM field outside the pumped active medium and may be evaluated as $d_{\text{beam}}^* \sim v_{\text{group}}/\gamma_a$.
For the parameters of the DFB laser we have $d_{\text{beam}}^* \sim 10 \mu \text{m}$, which is in agreement with the full model, Fig.~\ref{fig2}b.
Thus, because the delay time grows when $d_{\text{beam}}>d_{\text{beam}}^*$ and grows in the opposite case as well, we come to the conclusion that there is an optimal value, $\sim d_{\text{beam}}^*$, for the diameter of the pump beam for which the response time is a minimum.

\section*{Conclusion}

To summarize, we have studied the temporal dynamics of a two-dimensional plasmonic DFB laser under ultrashort pulse pumping.
Relying on numerical simulations and analytical calculations, we have shown that the response time of such a laser for typical experimental parameters can be decreased to one picosecond, which corresponds to a modulation frequency of $1$ THz.
This value is close to the upper limit for plasmonic lasers (the total response time can not be shorter than the time of the third stage, which is determined by the losses in the cavity, and is $ \sim 1$ ps).

It has been shown that the laser's response time can be shortened by optimizing the size of the pump beam, as a result of the consequent reduction of the delay time.
The dependence of the response time on the size of the pump beam is for following reason.
The interaction of the laser modes with the pumped active medium results in the emergence of a single collective mode.
If the size of the pump beam exceeds the propagation length of the EM waves, this collective mode is located in the pumped area.
The response time then grows logarithmically with increasing pump beam diameter.
In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.
Thus, there is an optimal size of the pump beam, for which the response time is a minimum.

The existence of an optimal size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
In turn, if the area of the pump beam is small, there is the opportunity to reduce the size of the device.
Thus, the existence of an optimal pump beam size paves the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.

\section*{Acknowledgements}

The reported study was funded by RFBR according to the research project No. 18-32-00596.

\bibliography{ref1}

\end{document}

### 2019-03-31 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Modified by AZ

@@
 	In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
-	We demonstrate that there is an optimal size of the pump beam for which the modulation speed is a maximum, and for typical experimental parameters can achieve $1$ THz.
+	We demonstrate that there is a critical minimal size of the pump beam for which the modulation speed is a maximum and can achieve $1$ THz.
 	The obtained result opens the way to increasing both the modulation speed and the energy efficiency of plasmonic laser devices.
@@
 This dependence is due to the complex structure of the modes of the DFB laser.
-The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump.
+The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump size.
 If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
-The response time in this case reaches picoseconds and slightly increases with the size of the pump beam.
+The response time in this case reaches picoseconds and almost does not change with the size of the pump beam.
 In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
-We demonstrate that there is an optimal size of the pump beam, for which the response time reaches a minimum.
-For typical experimental parameters, the optimal size of the pump beam is $\sim 15 \mu$m and the minimal response time is $\sim 1$ $\text{ps}$.
-The obtained value of the response time is in agreement with a recent experiment \cite{daskalakis2018nanolett} where it has been shown that plasmonic lasers have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime.
-Thus, the optimization of the pump beam size makes it possible to increase the modulation speed as well as the energy efficiency of plasmonic DFB laser devices.
+Therefore, there is a critical minimal diameter of the pump beam, $d_{beam}^*$, required for achievement of the short response time.
+This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
+When the size of pump beam is greater than $d_{beam}^*$ the response time of the plasmonic laser is $\sim 1$ $\text{ps}$.
+The obtained results confirm that plasmonic DFB lasers can have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime.
+The optimization of the pump beam size makes it possible to increase the modulation speed as well as the energy efficiency of plasmonic DFB laser devices.
 
@@
 $D_m$ is the population inversion of the $m$th atom and $\varphi _{jm} = - i \langle \hat{a} _j^{+} \hat{\sigma} _m \rangle$ describes the energy flow between the $m$th atom and $j$th mode.
-The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the eigenfrequencies of the modes; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms, $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
-$\Omega _{jm}$ is the Rabi constant of coupling between the $j$th mode and the $m$th atom.
-These equations generalize the rate equations widely used in laser theory \cite{SiegmanLasers}.
-In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the atoms, and the interference between the EM fields of the different modes, $n_{jk}$.
+The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the Bloch modes of plasmonic structure; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms, $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
+$\Omega _{jm}$ is the Rabi constant of coupling between the $j$th mode and the $m$th atom which can be expressed as
+\begin{equation}\label{eq4SM}
+{\Omega _{jm}} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar,
+\end{equation}
+where ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom.
+The electric field `per one quantum' is inversely proportional to the mode volume (see Appendix A).
+As a result, the Rabi constant increases with the decrease of the mode volume.
+The Eqs.~(\ref{eq1})--(\ref{eq3}) allow for the dependence of the stimulated and spontaneous emissions rates on the Rabi constant that enables to take into account the Purcell effect of the plasmonic structure.
+Every atom of the active medium has own coordinates which determine the Rabi constants between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
+That makes it possible to take into account the spatial distribution of the pumping and the population inversion within the active medium.
 
-For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes, see Fig.~\ref{fig1}.
-Such a structure has been realized in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
+The Eqs.~(\ref{eq1})--(\ref{eq3}) generalize the rate equations widely used in laser theory \cite{SiegmanLasers}.
+In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the atoms, and the interference between the EM fields of the different modes, $n_{jk}$ \cite{Zyablovsky2017approach}.
+
+For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes and coated the layer of the active medium, see Fig.~\ref{fig1}.
+We consider that the thickness of the Au film is $100$ $\text{nm}$, whereas the diameter of the holes is $160$ $\text{nm}$ and the distance between holes is $470$ $\text{nm}$.
+The surface of whole periodic plasmonic structure has size of 190 $\mu$m by 1908 $\mu$m.
+We consider that the transition frequency of the active medium is close to the frequency of second-order Bragg grating in the plasmonic structure and the linewidth of the active medium is about 100 nm.
+
+Similar plasmonic DFB laser have been realized in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
+
+In the absence of the holes the eigenmodes of the plasmonic structure are TM-modes similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
+These modes have in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of mode's propagation and both out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}.
+The in-plane electric field, $E_{\parallel}$, is in the direction of the mode's propagation. 
+The out-of-plane component of the electric field, $E_{\perp}$, is much greater than in-plane component, $E_{\parallel}$ \cite{TennerJOpt,TennerACSPhot}.
+The eigenmodes of the plasmonic structure with the holes we find within the coupled-mode theory with the experimental values of the scattering rates (see~\cite{TennerJOpt} and Section 2 in Suppl.Mat.
+of \cite{nefedkin2018acsphot}).
+The respective dispersion curves are shown \textbf{\underline{in Fig.~.}} The modes corresponding to different dispersion curves have various radiation losses and field distributions inside the unit cell of the plasmonic crystal (see Section 2 in Suppl.Mat.
+of \cite{nefedkin2018acsphot}).
+The far-field radiation from the plasmonic structure mainly arises owing to the scattering of $H_{\parallel}$ by the holes~\cite{TennerJOpt}.
+For this reason, the modes that have a minimum value of $H_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $H_{\parallel}$ at the holes.
+In the continuous-wave regime the lasing takes place at the modes with small radiation losses and high quality factor lied in 2nd dispersion curve, \textbf{\underline{see Fig.~}} \cite{TennerJOpt,nefedkin2018acsphot}.
+
+In numerical simulation we allow for several hundreds the Bloch modes of the plasmonic structure lied in different dispersion curves.
+The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat.
+of \cite{nefedkin2018acsphot}.
+
 We consider the response of the system to ultrashort pulse pumping of the active medium.
 Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}).
+We consider that in an initial moment the population inversions of the atoms whose coordinate lie inside the pump beam are expressed as
+\begin{equation}\label{eq5SM}
+{\left(D_m+1\right)/2} = {P_0 \Delta t / \left(n_c S h \hbar \omega \right)}. 
+\end{equation}
+Here $\left(D_m+1\right)/2$ is a population of the upper level of the atom; $P_0$ is a pump power; $\Delta t$ is a duration of the pump pulse; $n_c$ is a concentration of the active atoms; $S$ is an area of the pump beam; $h$ is the thickness of Au film; $\omega$ is a frequency of the pump pulse.
+Also, we consider that the population inversions of the atoms whose coordinate lie outside the pump beam are $-1$ in the initial moment.
 
@@
 	\end{subfigure}
-	\caption{Field distribution of the collective mode when area of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a) -(c) and $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f).
-The time moments in (a)-(c) and (d)-(f) correspond to circles on the red curve and the black curve in Fig.\ref{fig2}a, respectively.}
+	\caption{Distribution of the EM field intensity on the surface of plasmonic DFB laser when diameter of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a)-(c) in time moment 0.5 ps (a); 1.75 ps (b); 2.5 ps (c).
+Distribution of the EM field intensity on the surface of plasmonic DFB laser when diameter of the pump beam $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f) in time moment 0.32 ps (d); 0.541 ps (e); 1.5 ps (f).
+These time moments correspond to circles on the red curve and the black curve in Fig.\ref{fig2}a, respectively.
+The dashed circles in Figs.
+3 denote the boundaries of the pump beam.}
 	\label{fig3}	
@@
 The existence of an optimal size of the pump beam can be explained with the example of a single-mode model.
-To justify the single-mode approximation, we calculate the distribution of the total electric field over the area of the DFB laser at different moments of time, see Fig.~\ref{fig3}.
+To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
+Although, in our simulation we take into account the several hundreds of the Bloch modes of the plasmonic structure which jointly form the EM field distribution in the active medium we can see that the EM field distribution practically does not change in the first and second stage of the system response.
+Another words, there is dynamical balance between the relaxation processed in the plasmonic structure, the propagation of the EM field and the stimulated amplification of the EM field in the active medium.
 When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
-In the opposite case, $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
-However, in both cases, the spatial distribution of the total electric field remains almost unchanged during its temporal evolution.
-The time independence of the EM field distribution allows us to consider the plasmonic laser as a single mode laser.
+Whereas when $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
+However, in both cases, the spatial distribution of the electric field intensity remains almost unchanged during its temporal evolution.
 
-Based on this observation, we assume that the part of the active medium pumped by an external beam effectively excites a single collective mode, which is formed by the Bloch modes of the plasmonic structure.
-In this case, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
+We use the fact of the constancy of the EM field distribution for construction of the single-mode model.
+For this purpose, we consider that there is single collective mode whose the EM field distribution coincide with the EM field distribution found in the numerical simulation multi-mode problem.
+Furthermore, we assume that the pumped active medium interacts only with such a single collective mode.
+Within of these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
 \begin{gather} 
@@
 
-Note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
-The superradiance problem and the problem that we investigate here have the same physical grounds.
-Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).
-
 The figure \ref{fig5}b shows that the analytical approximation, Eq.~(\ref{tdel}), agrees well with the numerical simulation of the exact model, Eqs. (\ref{eq1})--(\ref{eq3}).
@@
 In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.
-Thus, there is an optimal size of the pump beam, for which the response time is a minimum.
 
-The existence of an optimal size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
+Thus, there is a critical minimal diameter of the pump beam, $d_{beam}^*$, required for achievement of the short response time.
+This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
+The existence of the critical pump beam size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
 In turn, if the area of the pump beam is small, there is the opportunity to reduce the size of the device.
-Thus, the existence of an optimal pump beam size paves the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.
+Thus, the obtained results pave the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.
+
+It is interesting that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
+Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).
+In the plasmonic cavity the inequality $\gamma_a \gg \gamma_D$ always is hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds.
+We believe that further study of connection between these problems is interesting.
 
@@
 
+\section*{Appendix A}
+To normalize the electric field ${{\bf{E}}_j}({{\bf{r}}_m})$, we equate the energy of one quantum to the energy of the electric field in the resonator:
+\begin{equation}
+\hbar {\omega _j} = \frac{1}{{8\pi }}\int\limits_V {\left( {{{\left.
+{\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _j}}}{{\left| {{{\bf{E}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{H}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}} \tag{A.1}
+\end{equation}
+where $\varepsilon (\omega, \textbf{r})$  is the dielectric constant of the active plasmonic structure.
+
+In a two-dimensional periodic plasmonic structure whose surface is parallel to the $xy$-plane, the electric and magnetic fields satisfy the Bloch condition,
+\begin{equation}
+{{\bf{E}}_j}\left( {\bf{r}} \right) = {A_{0j}}\exp \left( {i{k_{xj}}x + i{k_{yj}}y} \right){{\bf{e}}_j}\left( {x,y,z} \right) \tag{A.2}
+\end{equation}
+\begin{equation}
+{{\bf{H}}_j}\left( {\bf{r}} \right) = {A_{0j}}\exp \left( {i{k_{xj}}x + i{k_{yj}}y} \right){{\bf{h}}_j}\left( {x,y,z} \right), \tag{A.3}
+\end{equation}
+where $A_{0j}$ is a real normalization constant for \textit{j}th mode; ${{\bf{e}}_j}\left( {x,y,z} \right) = {{\bf{e}}_j}\left( {x + {L_x},y + {L_y},z} \right)$  and ${{\bf{h}}_j}\left( {x,y,z} \right) = {{\bf{h}}_j}\left( {x + {L_x},y + {L_y},z} \right)$, where $L_x$ and $L_y$ are periods of the plasmonic structure along $x$-axis and $y$-axis, respectively.
+
+Suppose that the area under investigation contains an integral number $N$ of plasmonic lattice cells.
+Then Eq. (A.1) may be rewritten as
+\begin{equation}
+\hbar {\omega _j} =\frac{N}{{8\pi }} A_{0j}^2 \int\limits_{{V_{cell}}} {\left( {{{\left.
+{\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}} \tag{A.4}
+\end{equation}
+where ${V_{cell}}$  is the volume of the cell of the plasmonic lattice. As a result, we obtain that the normalization constant $A_{0j}$ is expressed as
+\begin{equation}
+A_{0j} =\sqrt{\frac{\hbar {\omega _j}}{\frac{N}{{8\pi }} \int\limits_{{V_{cell}}} {\left( {{{\left.
+{\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}}}} \tag{A.5}
+\end{equation}
+
+The Rabi constant between \textit{j}th mode and \textit{m}th atom, ${\Omega _{jm}}$, is expressed as
+\begin{equation}
+{\Omega _{jm}} =  - {{{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar} \tag{A.6}
+\end{equation}
+where ${\bf{d}}_m$ is a dipole moment of the \textit{m}th atom of the active medium.
+
+From Eqns. (A.2), and (A.7) we obtain that
+\begin{equation}
+{\Omega _{jm}} =-\sqrt{\frac{{8\pi \omega _j}}{{N \hbar} \int\limits_{{V_{cell}}} {\left( {{{\left.
+{\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}}}} \left( {{{\bf{d}}_m} \cdot {{\bf{e}}_j}\left( {{{\bf{r}}_m}} \right)} \right)\exp \left( {i{k_{xj}}{x_m} + i{k_{yj}}{y_m}} \right) \tag{A.7}
+\end{equation}
+Thus, the Rabi constant may be rewritten as
+\begin{equation}
+{\Omega _{jm}} = {\Omega _{0jm}}\exp \left( {i{k_{xj}}{x_m} + i{k_{yj}}{y_m}} \right){f_j}\left( {{{\bf{r}}_m}} \right) \tag{A.8}
+\end{equation}
+where $\Omega _{0jm}$ can be expressed as
+\begin{equation}
+{\Omega _{0jm}} =-\sqrt{\frac{{8\pi \omega _j \left|\textbf{d}_m\right|^2 \left|\textbf{e}_{j}\right|_{max}^2}}{{N \hbar} \int\limits_{{V_{cell}}} {\left( {{{\left.
+{\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}}}} = -\sqrt{\frac{{8\pi \omega _j \left|\textbf{d}_m\right|^2}}{\hbar V_{j mode}}} \tag{A.9}
+\end{equation}
+where $\left|\textbf{e}_{j}\right|_{max}$ is a maximum value of the periodic function $\left|\textbf{e}_{j}(\textbf{r})\right|$; $V_{j mode}=\frac{\int\limits_{{V_{cell}}} {\left( {{{\left.
+{\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}}}{\left|\textbf{e}_{j}\right|_{max}^2}$ is mode volume of \textit{j}th mode.
+Also, ${f_j}\left( {{{\bf{r}}_m}} \right) = \left( {{{\bf{d}}_m} \cdot {{\bf{e}}_j}\left( {{{\bf{r}}_m}} \right)} \right)/(\left|\bf{d}_m\right| \left|\textbf{e}_{j}\right|_{max})$ is a periodic function with plasmonic lattice cell period.
+Note that the functions ${f_j}\left( {{{\bf{r}}_m}} \right)$ for the modes lied in different bands are orthogonal to each other.
+The orthogonality arises due to difference of the EM field distributions within the cell of the plasmonic structure for the modes lied in different bands.
+
+\section*{Appendix B}
+The eigenimodes of such a plasmonic structure without holes are the TM-modes that are similar to surface plasmons at the metal-dielectric interface.
+These modes have in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of the mode propagation, and out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field.
+The in-plane electric field, $E_{\parallel}$, is in the direction of the mode propagation.
+For the surface plasmons the out-of-plane component of the electric field, $E_{\perp}$, is much greater than the in-plane component, $E_{\parallel}$.
+
+In a periodical lattice, the electric and magnetic fields of the eigenmodes satisfy the Bloch condition. 
+We consider that the transition frequency of the active medium is close to the frequency at which the second Bloch condition is satisfied, i.e.
+$k_{sp}(\omega_{\sigma}) \approx G$, where $k_{sp}$ is a wavevector of the TM-mode of the plasmonic structure without holes and $G$ is a reciprocal lattice vector.
+The eigenmodes of the 2D plasmonic DFB laser can be calculated by the coupled-mode theory~\cite{ExterOptExp,TennerJOpt}. 
+Within this framework the electric field is expressed as
+%\begin{widetext}
+\begin{equation}
+\begin{array}{l}
+{\bf{E}}\left( {r,t} \right) = \exp \left( {i{k_x}x + i{k_y}y} \right) \times \\
+\times \left( {{E_x}\left( t \right){{\bf{u}}_x}\exp \left( {iGx} \right) + {E_{ - x}}\left( t \right){{\bf{u}}_{ - x}}\exp \left( { - iGx} \right) + {E_y}\left( t \right){{\bf{u}}_y}\exp \left( {iGy} \right) + {E_{ - y}}\left( t \right){{\bf{u}}_{ - y}}\exp \left( { - iGy} \right)} \right)\tag{5}
+\end{array}
+\end{equation}
+%\end{widetext}
+Here, we suppose that the surface of the Au film lies perpendicularly to z axis and the sides of the plasmonic lattice are in the direction of x and y axes. 
+The centre of the unit cell coincides with the centre of a hole.
+The vectors ${{\bf{u}}_x}$,  ${{\bf{u}}_{ - x}}$, ${{\bf{u}}_y}$, and ${{\bf{u}}_{ - y}}$  are the eigenvectors of propagating TM-modes of the plasmonic system without holes with respectively directed wavevectors.
+${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$  are the amplitudes of the corresponding eigenmodes and ${k_{x,y}} = \omega /c\sin {\theta _{x,y}}$, where ${\theta _{x,y}}$ is the emission angle along the respective axes.
+
+The amplitudes  ${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$ are found from
+\begin{equation}
+dE/dt = i\hat WE = i\left( {{{\hat W}_0} + \hat V} \right)E\tag{6}
+\end{equation}
+where 
+%\begin{strip}
+\begin{equation}
+\begin{array}{l}
+{\hat W_0} = \left( {\begin{array}{*{20}{c}}
+	{{\omega _0}}&\gamma &\kappa &\kappa \\
+	\gamma &{{\omega _0}}&\kappa &\kappa \\
+	\kappa &\kappa &{{\omega _0}}&\gamma \\
+	\kappa &\kappa &\gamma &{{\omega _0}}
+	\end{array}} \right) , \\
+\hat V\left( {{\theta _x},{\theta _y}} \right) = \left( {\begin{array}{*{20}{c}}
+	{{c_2}\theta _x^2 + {c_1}{\theta _y}}&0&0&0\\
+	0&{{c_2}\theta _x^2 - {c_1}{\theta _y}}&0&0\\
+	0&0&{{c_1}{\theta _x} + {c_2}\theta _y^2}&0\\
+	0&0&0&{ - {c_1}{\theta _x} + {c_2}\theta _y^2}
+	\end{array}} \right)\tag{7}
+\end{array}
+\end{equation}
+%\end{strip}
+Here, ${\omega _0} = Gc/{n_{eff}}$ ($c$ is the speed of light), $\gamma $  and $\kappa$  are the back-scattering amplitude at angle $180^{\circ }$ and the right-angle scattering at $\pm 90^{\circ }$, respectively; ${c_1} = {\omega _0}/{n_{eff}}$, ${c_2} = {\omega _0}/\left( {2n_{eff}^2} \right)$, and ${n_{eff}}$  is an effective refractive index~\cite{ExterOptExp,TennerJOpt}.
+The eigenvectors of $\hat H$  estimate the amplitudes ${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$.
+The values of $\gamma $, $\kappa$, $c_1$, $c_2$ and $n_{eff}$ are taken from~\cite{TennerJOpt}.
+Using them, we obtain the dispersion curves for the 2D distributed-feedback laser, see Figure 2 in the manuscript.
+
+At $\theta _x = \theta _y = 0$ the eigenvectors of $\hat W$ are \text{(1,1,1,1)}, \text{(1,1,-1,-1)}, \text{(0,0,1,-1)} and \text{(1,-1,0,0)}. 
+These eigenvectors correspond to the following electric field distributions:
+
+\textbf{Table S1.}
+
+$\begin{array}{cccc}
+\text{The eigenvector} & E_{\parallel} & \
+H_{\parallel} & \
+E_ {\perp } \left( E_z \right) \\
+\text{(1,1,1,1)} & \sin (G x)+\sin (G y) & \sin (G x)+\sin (G y) & \
+\cos (G x)+\cos (G y) \\
+\text{(1,1,-1,-1)} & \sin (G x)-\sin (G y) & \sin (G x)-\sin (G y) & \
+\cos (G x)-\cos (G y) \\
+\text{(0,0,1,-1)} & \cos (G y) & \cos (G y) & \sin (G y) \\
+\text{(1,-1,0,0)} & \cos (G x) & \cos (G x) & \sin (G x) \\
+\end{array}$
+
+The dispersion curves in the main text are numbered in the same order as in Table 1.
+The modes radiate the far-fields because of the scattering of electromagnetic waves by the holes~\cite{TennerJOpt}. 
+The TM-modes dominantly couple with far-field via the in-plane magnetic field in the holes.
+Therefore, the first and second modes which have minimum of the in-plane magnetic field in the centre of the unit cell of the plasmonic lattice (see Table 1) do not radiate, i.e. they are dark modes. 
+On the contrary, the third and fourth modes are bright modes and they radiate.
+
 \bibliography{ref1}

### 2019-04-01 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Modified by NN

@@
 	In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
-	We demonstrate that there is a critical minimal size of the pump beam for which the modulation speed is a maximum and can achieve $1$ THz.
+	We demonstrate that there is a critical minimum size of the pump beam for which the modulation speed is a maximum and can achieve $1$ THz.
 	The obtained result opens the way to increasing both the modulation speed and the energy efficiency of plasmonic laser devices.
@@
 If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
-The response time in this case reaches picoseconds and almost does not change with the size of the pump beam.
+The response time in this case reaches picoseconds and practically does not change with the size of the pump beam.
 In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
-Therefore, there is a critical minimal diameter of the pump beam, $d_{beam}^*$, required for achievement of the short response time.
-This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
-When the size of pump beam is greater than $d_{beam}^*$ the response time of the plasmonic laser is $\sim 1$ $\text{ps}$.
+Consequently, there is a critical minimum diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the shortest response time.
+This diameter is about the decay length of the EM field in the plasmonic structure and for the plasmonic laser under consideration is $\sim 15 \mu$m.
+When the size of the pump beam is greater than $d_{\text{beam}}^*$, the response time of the plasmonic laser is $\sim 1$ $\text{ps}$.
 The obtained results confirm that plasmonic DFB lasers can have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime. 
-The optimization of the pump beam size makes it possible to increase the modulation speed as well as the energy efficiency of plasmonic DFB laser devices.
+Optimization of the size of the pump beam allows an increase in the modulation speed, as well as the energy efficiency of plasmonic DFB laser devices.
 
@@
 $D_m$ is the population inversion of the $m$th atom and $\varphi _{jm} = - i \langle \hat{a} _j^{+} \hat{\sigma} _m \rangle$ describes the energy flow between the $m$th atom and $j$th mode.
-The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the Bloch modes of plasmonic structure; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms, $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
+The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the Bloch modes of plasmonic structure; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms; $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
 $\Omega _{jm}$ is the Rabi constant of coupling between the $j$th mode and the $m$th atom which can be expressed as
@@
 The Eqs.~(\ref{eq1})--(\ref{eq3}) allow for the dependence of the stimulated and spontaneous emissions rates on the Rabi constant that enables to take into account the Purcell effect of the plasmonic structure.
-Every atom of the active medium has own coordinates which determine the Rabi constants between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
-That makes it possible to take into account the spatial distribution of the pumping and the population inversion within the active medium.
+Every atom of the active medium has its coordinates, which determine the Rabi constants between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
+This allows us to take into account the spatial distribution of the pumping and the population inversion within the active medium.
 
@@
 
-For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes and coated the layer of the active medium, see Fig.~\ref{fig1}.
-We consider that the thickness of the Au film is $100$ $\text{nm}$, whereas the diameter of the holes is $160$ $\text{nm}$ and the distance between holes is $470$ $\text{nm}$.
-The surface of whole periodic plasmonic structure has size of 190 $\mu$m by 1908 $\mu$m.
-We consider that the transition frequency of the active medium is close to the frequency of second-order Bragg grating in the plasmonic structure and the linewidth of the active medium is about 100 nm.
+For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes and coated with the layer of the active medium, see Fig.~\ref{fig1}.
+We consider that the thickness of the Au film is $100$ $\text{nm}$, the diameter of the holes is $160$ $\text{nm}$, and the distance between holes is $470$ $\text{nm}$.
+The surface of the entire periodic plasmonic structure is $190 \times 190$ $\mu \text{m}^2$.
+We consider that the transition frequency of the active medium is close to the frequency of the second-order Bragg grating in the plasmonic structure and the linewidth of the active medium is about 100 nm.
 
-Similar plasmonic DFB laser have been realized in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
+A similar plasmonic DFB laser have been implemented in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
 
-In the absence of the holes the eigenmodes of the plasmonic structure are TM-modes similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
-These modes have in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of mode's propagation and both out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}.
+In the absence of the holes, the eigenmodes of the plasmonic structure are TM-modes, similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
+These modes have an in-plane component of the magnetic field, $H_{\parallel}$, which is perpendicular to the direction of mode's propagation and both out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}.
 The in-plane electric field, $E_{\parallel}$, is in the direction of the mode's propagation. 
-The out-of-plane component of the electric field, $E_{\perp}$, is much greater than in-plane component, $E_{\parallel}$ \cite{TennerJOpt,TennerACSPhot}.
-The eigenmodes of the plasmonic structure with the holes we find within the coupled-mode theory with the experimental values of the scattering rates (see~\cite{TennerJOpt} and Section 2 in Suppl.Mat.
+The out-of-plane component of the electric field, $E_{\perp}$, is much greater than the in-plane component, $E_{\parallel}$ \cite{TennerJOpt,TennerACSPhot}.
+We find the eigenmodes of the plasmonic structure with the holes within the coupled-mode theory with the experimental values of the scattering rates (see~\cite{TennerJOpt} and Section 2 in Suppl.Mat.
 of \cite{nefedkin2018acsphot}).
-The respective dispersion curves are shown \textbf{\underline{in Fig.~.}} The modes corresponding to different dispersion curves have various radiation losses and field distributions inside the unit cell of the plasmonic crystal (see Section 2 in Suppl.Mat.
+Corresponding dispersion curves are shown in \textbf{\underline{in Fig.~.}} The modes corresponding to different dispersion curves have various radiation losses and field distributions inside the unit cell of the plasmonic crystal (see Section 2 in Suppl.Mat.
 of \cite{nefedkin2018acsphot}).
-The far-field radiation from the plasmonic structure mainly arises owing to the scattering of $H_{\parallel}$ by the holes~\cite{TennerJOpt}.
+The far-field radiation from the plasmonic structure arises mainly due to the scattering of $H_{\parallel}$ on the holes~\cite{TennerJOpt}.
 For this reason, the modes that have a minimum value of $H_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $H_{\parallel}$ at the holes.
-In the continuous-wave regime the lasing takes place at the modes with small radiation losses and high quality factor lied in 2nd dispersion curve, \textbf{\underline{see Fig.~}} \cite{TennerJOpt,nefedkin2018acsphot}.
+In the continuous-wave regime, the lasing takes place at the modes with small radiation losses and a high quality factor lying on the 2nd dispersion curve, \textbf{\underline{see Fig.~}} \cite{TennerJOpt,nefedkin2018acsphot}.
 
-In numerical simulation we allow for several hundreds the Bloch modes of the plasmonic structure lied in different dispersion curves.
+In numerical simulations, we take into account several hundreds of the Bloch modes of the plasmonic structure lying on different dispersion curves.
 The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat.
@@
 Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}).
-We consider that in an initial moment the population inversions of the atoms whose coordinate lie inside the pump beam are expressed as
+We consider that at an initial time the population inversions of the atoms whose coordinates lie inside the pump beam are expressed as
 \begin{equation}\label{eq5SM}
@@
 Here $\left(D_m+1\right)/2$ is a population of the upper level of the atom; $P_0$ is a pump power; $\Delta t$ is a duration of the pump pulse; $n_c$ is a concentration of the active atoms; $S$ is an area of the pump beam; $h$ is the thickness of Au film; $\omega$ is a frequency of the pump pulse.
-Also, we consider that the population inversions of the atoms whose coordinate lie outside the pump beam are $-1$ in the initial moment.
+Also, we consider that the population inversions of the atoms whose coordinates lie outside the pump beam are $-1$ at the initial time.
 
@@
 	\end{subfigure}
-	\caption{Distribution of the EM field intensity on the surface of plasmonic DFB laser when diameter of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a)-(c) in time moment 0.5 ps (a); 1.75 ps (b); 2.5 ps (c).
-Distribution of the EM field intensity on the surface of plasmonic DFB laser when diameter of the pump beam $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f) in time moment 0.32 ps (d); 0.541 ps (e); 1.5 ps (f).
+	\caption{Distribution of the EM field intensity on the surface of the plasmonic DFB laser when the diameter of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a)-(c) in time moment 0.5 ps (a); 1.75 ps (b); 2.5 ps (c).
+Distribution of the EM field intensity on the surface of the plasmonic DFB laser when the diameter of the pump beam $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f) in time moment 0.32 ps (d); 0.541 ps (e); 1.5 ps (f).
 These time moments correspond to circles on the red curve and the black curve in Fig.\ref{fig2}a, respectively.
-The dashed circles in Figs.
-3 denote the boundaries of the pump beam.}
+The dashed circles denote the boundaries of the pump beam.}
 	\label{fig3}	
@@
 To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
-Although, in our simulation we take into account the several hundreds of the Bloch modes of the plasmonic structure which jointly form the EM field distribution in the active medium we can see that the EM field distribution practically does not change in the first and second stage of the system response.
-Another words, there is dynamical balance between the relaxation processed in the plasmonic structure, the propagation of the EM field and the stimulated amplification of the EM field in the active medium.
+Although in our simulations we take into account several hundreds of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response.
+In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field and the stimulated amplification of the EM field in the active medium.
 When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
-Whereas when $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
+In the opposite case, when $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
 However, in both cases, the spatial distribution of the electric field intensity remains almost unchanged during its temporal evolution.
 
-We use the fact of the constancy of the EM field distribution for construction of the single-mode model.
-For this purpose, we consider that there is single collective mode whose the EM field distribution coincide with the EM field distribution found in the numerical simulation multi-mode problem.
+We use the fact of the constancy of the EM field distribution to construct a single-mode model.
+To do this, we assume that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multi-mode problem.
 Furthermore, we assume that the pumped active medium interacts only with such a single collective mode.
-Within of these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
+Within these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
 \begin{gather} 
@@
 
-Thus, there is a critical minimal diameter of the pump beam, $d_{beam}^*$, required for achievement of the short response time.
+Thus, there is a critical minimum diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the shortest response time.
 This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
@@
 
-It is interesting that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
+It is interesting to note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
 Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).
-In the plasmonic cavity the inequality $\gamma_a \gg \gamma_D$ always is hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds.
-We believe that further study of connection between these problems is interesting.
+In the plasmonic cavity, the inequality $\gamma_a \gg \gamma_D$ is always hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds.
+We believe that further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.
 
@@
 Also, ${f_j}\left( {{{\bf{r}}_m}} \right) = \left( {{{\bf{d}}_m} \cdot {{\bf{e}}_j}\left( {{{\bf{r}}_m}} \right)} \right)/(\left|\bf{d}_m\right| \left|\textbf{e}_{j}\right|_{max})$ is a periodic function with plasmonic lattice cell period.
-Note that the functions ${f_j}\left( {{{\bf{r}}_m}} \right)$ for the modes lied in different bands are orthogonal to each other.
-The orthogonality arises due to difference of the EM field distributions within the cell of the plasmonic structure for the modes lied in different bands.
+Note that the functions ${f_j}\left( {{{\bf{r}}_m}} \right)$ for the modes lying in different bands are orthogonal to each other.
+The orthogonality arises due to difference of the EM field distributions within the cell of the plasmonic structure for the modes lying in different bands.
 
@@
 The eigenimodes of such a plasmonic structure without holes are the TM-modes that are similar to surface plasmons at the metal-dielectric interface.
-These modes have in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of the mode propagation, and out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field.
+These modes have an in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of the mode propagation, and out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field.
 The in-plane electric field, $E_{\parallel}$, is in the direction of the mode propagation.
@@
 
-In a periodical lattice, the electric and magnetic fields of the eigenmodes satisfy the Bloch condition. 
+In a periodic lattice, the electric and magnetic fields of the eigenmodes satisfy the Bloch condition. 
 We consider that the transition frequency of the active medium is close to the frequency at which the second Bloch condition is satisfied, i.e.
@@
 %\end{widetext}
-Here, we suppose that the surface of the Au film lies perpendicularly to z axis and the sides of the plasmonic lattice are in the direction of x and y axes. 
+Here, we suppose that the surface of the Au film lies perpendicularly to $z$ axis and the sides of the plasmonic lattice are in the direction of $x$ and $y$ axes. 
 The centre of the unit cell coincides with the centre of a hole.
 The vectors ${{\bf{u}}_x}$,  ${{\bf{u}}_{ - x}}$, ${{\bf{u}}_y}$, and ${{\bf{u}}_{ - y}}$  are the eigenvectors of propagating TM-modes of the plasmonic system without holes with respectively directed wavevectors.
-${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$  are the amplitudes of the corresponding eigenmodes and ${k_{x,y}} = \omega /c\sin {\theta _{x,y}}$, where ${\theta _{x,y}}$ is the emission angle along the respective axes.
+${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$  are the amplitudes of the corresponding eigenmodes, and ${k_{x,y}} = \omega /c\sin {\theta _{x,y}}$, where ${\theta _{x,y}}$ is the angle of emission along the respective axes.
 
@@
 %\end{strip}
-Here, ${\omega _0} = Gc/{n_{eff}}$ ($c$ is the speed of light), $\gamma $  and $\kappa$  are the back-scattering amplitude at angle $180^{\circ }$ and the right-angle scattering at $\pm 90^{\circ }$, respectively; ${c_1} = {\omega _0}/{n_{eff}}$, ${c_2} = {\omega _0}/\left( {2n_{eff}^2} \right)$, and ${n_{eff}}$  is an effective refractive index~\cite{ExterOptExp,TennerJOpt}.
+Here, ${\omega _0} = Gc/{n_\text{eff}}$ ($c$ is the speed of light), $\gamma $  and $\kappa$  are the back-scattering amplitude at angle $180^{\circ }$ and the right-angle scattering at $\pm 90^{\circ }$, respectively; ${c_1} = {\omega _0}/{n_\text{eff}}$, ${c_2} = {\omega _0}/\left( {2n_\text{eff}^2} \right)$, and ${n_\text{eff}}$  is an effective refractive index~\cite{ExterOptExp,TennerJOpt}.
 The eigenvectors of $\hat H$  estimate the amplitudes ${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$.
-The values of $\gamma $, $\kappa$, $c_1$, $c_2$ and $n_{eff}$ are taken from~\cite{TennerJOpt}.
-Using them, we obtain the dispersion curves for the 2D distributed-feedback laser, see Figure 2 in the manuscript.
+The values of $\gamma $, $\kappa$, $c_1$, $c_2$ and $n_\text{eff}$ are taken from~\cite{TennerJOpt}.
+Using them, we obtain the dispersion curves for the 2D DFB laser, see \textbf{Figure 2}.
 

### 2019-04-01 Response_DFB_Pulse_PRApplied.md
Response modified by NN

----------------------------------------------------------------------

# Report of Referee A -- NB10105/Nefedkin

----------------------------------------------------------------------

*This paper theoretically studied the ultrafast response time of a plasmonic distributed feedback laser in the large-signal modulation regime. The authors demonstrated the dependence of response limit with*
*pump spot size using both numerical and analytical approaches. The identification on different ultrafast stages and insights on pump-tunable lasing dynamics could facilitate the applications of plasmon lasers in optoelectronic devices.*

**Review**

*However, the manuscript focused on plasmonic nanohole arrays with a specific parameter design (100-nm hole, 470-nm spacing and 100-nm film thickness). Are the conclusions on the critical pump spot size and*
*ultrafast response limit (1 ps) general enough to plasmon nanohole array lasers in different design parameters (film thickness, hole size, spacing)?*

**Response**

**1.** The time response of the plasmonic laser is determined by the Purcell factor and the EM field relaxation rate in plasmonic structure. These characteristics depend on film thickness, hole size, spacing. For
this reason, the values of the critical pump spot size and ultrafast response limit are different for divergent realizations of plasmon nanohole array lasers.

However, our analytical model shows that the existence of the critical pump spot size is related to finite size of an effective laser mode. When the size of pump beam is greater than the size of the effective
laser mode the response time almost does not depend on the size of pump beam. When the size of the pump beam becomes smaller than the size of the effective laser mode, the response time increases with the decrease of the pump beam size. For this reason, the critical pump spot size takes place in different realizations of the plasmonic distributed-feedback lasers.

**Review**

*Also, plasmon lasers based on nanohole arrays and nanoparticle arrays show different lasing mechanisms from surface plasmon polariton and surface lattice resonances, respectively. It is not a fair comparison*
*between the calculated response time of plasmon nanohole arrays and experimental results of nanoparticle arrays to verify the accuracy of theoretical modeling (final paragraph, the introduction section).*

**Response**

**2.** It is true that in our numerical simulations we consider plasmon lasers based on nanohole arrays, though in the work [Torma] the experiment has been conducted for plasmon lasers based on nanoparticle arrays in the different geometry. In this regard, the conformity of the response times obtained in the experiment and in our simulations is not a fair verification of the accuracy of our theoretical modeling. We agree with the referee in this point and have reformulated the last paragraph in the Introduction section by excluding this direct comparison.

**Review**

*In addition, can the authors more explanations and rationales on how experimental conditions such as pump size and pump power density (excitation fluence) are incorporated in the modeling?*

**Response**

**3.** In our modeling we take into account a large amount of atoms, which have different coordinates $\textbf{r}_m$. The coupling constant between $j$th mode of the plasmonic structure and $m$th atom depends on atom's coordinates as
$$
\Omega_{jm} = \Omega_0 e^{i \textbf{k}_j \textbf{r}_m}
\tag{1}
$$

The size of the pump beam and the power density of pumping are incorporated in the model as the initial conditions for population inversions of atoms. The initial value of the population inversion for the atoms located inside the pumping spot is $D_0$. Other atoms have initial population inversions equal to $-1$, i.e. they are in the ground state. The value of $D_0$ is obtained from the pumping power density as
$$
\frac{D_0 + 1}{2} = \frac{P_0 \Delta t}{n_c S l \hbar \omega}
\tag{2}
$$
Here $P_0$ is power of the laser pulse pump; $\Delta t$ is duration of the pump pulse; $n_c$ is the concentration of atoms of the active medium; $S$ is the area of the pump beam; $l$ is the width of the active medium layer; $\hbar \omega$ is the transition frequency of the atom; and $(D_0 + 1)/2$ is the population of the excited state arising due to the pump pulse.

Thus, we assume that the entire pulse energy comes to the excitation of atoms of the active medium. In real experiment, there are additional channels of energy losses, which lead to the decrease of the population inversion of the excited state. These losses can be taken into account as an addition factor in (2). For the simplicity, we do not consider the additional losses and use the estimation (2).

In our simulations the power of the pump pulse is $P_0 \approx $. This value is typical for experiments carried out in with the same plasmonic structure in the same geometry, see [van exter, Beijnum].

Note that the dependence of the response time on the pump beam size is observable in a wide range of $D_0$ values, and a fine tune of the pumping parameters is not necessary to observe the predicted dependency.

**Review**

*Did the authors consider any diffractive coupling between nanoholes in the equations of system dynamics, given that pump size may influence the resonance mode quality (Purcell factor) and EM fields near the holes?*

**Response**

**4.** We take into account diffractive coupling between nanoholes in laser equations. For this purpose, firstly, we find the EM field distribution of the Bloch modes of the plasmonic structure within the coupled-mode theory. Using this framework we allow for the scattering of the surface plasmon-polariton waves with different Bloch wavenumbers on the nanoholes. We use the scattering rates found in Ref. \[V. T. Tenner, A. N. van Delft, M. J. A. de Dood, and M. P. van Exter, *Journal of Optics* 16, 114019 (2014)\] for the plasmonic laser based on Au film perforated by the nanohole arrays. The coupled-mode theory with such scattering rates demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser \[V. T. Tenner, A. N. van Delft, M. J. A. de Dood, and M. P. van Exter, Journal of Optics 16, 114019 (2014)\].

Secondly, we use the obtained EM field distributions to calculate the Rabi constants between the Bloch modes and atoms of the active medium. The eigenfrequencies and the relaxation rates of the Bloch modes enter in the laser equations (1)-(3).

Thus, the parameters and dispersive dependences of the Bloch modes of the plasmonic structure taken into account in the laser equations via the Rabi constants, the eiegnfrequencies, and the relaxation rates.

We added the description of the band-structure of the plasmonic nanoholes array laser in the Section 2 of main part of the manuscript and **<u>Appendixes A and B.</u>**

A change in the pump beam size results in a change in the volume of the collective mode. As a result, the Purcell factor and the rates of the spontaneous and stimulated emissions also change. The strongest change occurs when the diameter of the pump beam is less than the critical value, at which the strongest dependence of the response time on the diameter appears.

Note that in our modeling the diameter of the pump beam is always much larger than the period of the nanohole array. For this reason, we do not take into account the inhomogeneous pumping within the cell of plasmonic structure.

**Review**

*Also, how is the Rabi constant in the modeling related to the Purcell factor of a specific plasmon cavity? More justifications are needed.*

**5.** We added the derivation of the Rabi constant through the parameters of the plasmonic structure and the active medium in the revised version of the manuscript , see Section 2 in main part of the text and the Appendix A.

**Review**

*We have other comments below*

**1.** *On page 4, the authors claimed that “The time independence of the EM field distribution allows us to consider the plasmonic laser as a single mode laser,” but they need to justify how the time evolution of EM field is correlated to single-mode approximation. Is there any effect from high-order diffractive modes for the hole arrays? Why does the EM field distribution slightly change from Fig 3d to 3f?*

**Response**

In our simulations we take into account several hundred of the Bloch modes of the plasmonic structure which form together the EM field distribution in the active medium. However, our simulations show that the resulting the EM field distribution slightly changes in the first and second stage of the output pulse evolution. In other words, there is a dynamical equilibrium between relaxation and propagation of the EM field on the one hand and the stimulated amplification of the EM field in the active medium on the other hand. We use the fact of the constancy of the EM field distribution to construct the single-mode model. We consider that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multimode problem. This assumption allows to derive the single-mode laser equations. The results obtained in the framework of the single-mode model are in good agreement with the results of the multimode model, see Figs. 2(a), 4 and Fig. 5a.

The change of the EM field distribution from Fig 3d to 3f is associated with a stronger decrease of the population inversion in the third stage of the response. This leads to a violation of  the dynamical equilibrium between relaxation, propagation of the EM field and the stimulated amplification of the EM field in the system. As a result, the EM field distribution begins to change, and the intensity of the EM field begins to decrease.

We added the appropriate explanations in the revised version of the manuscript.

**2.** *Regarding the figures, the order of Figure 5b and 5a can be switched since 5b appeared earlier in the main text. In Figure 5b, what are the units of t_delay? Figure 3 needs additional labeling such as scale bar and timing point of chosen snapshots, and the authors can add more explanations in the figure caption such as the meaning of dashed circles.*

**Response**

We have made the required changes. We changed the order of Figs. 5a and 5b. We indicated the units for $t_\text{delay}$ in Fig. 5b. We added scale bar to the Figure 3; we indicated in the figure caption the time
moments corresponding to the intensity field distribution in Figures 3. We described the meaning of dashed circles in the Figure 3.

We also changed the caption to Figure 3.

In the previous version:

"Field distribution of the collective mode when ..."

In a new version:

"Distribution of the EM field intensity on the surface of plasmonic DFB
laser when ..."

**3.** *The authors point out the similarity of Eq. (15) to the time dependence of a superradiant burst, and how that system resembles the current problem and approach (page 6). The parallelism between these two systems is interesting; however, these few sentences seem out of place relative to the major discussion on ultrafast stages of the laser response. We suggest this information be moved to the introduction or conclusion section, and presented as either as potential motivation or future prospect of study.*

**Response**

We placed all mentions to the superradiance problem to the Conclusion section of the manuscript and noted the potential interest in studying the similarity of the observed phenomenon and superradiance.

----------------------------------------------------------------------

# Report of Referee B -- NB10105/Nefedkin

----------------------------------------------------------------------

*In this manuscript, the Authors theoretically analyze the dynamics of a two-dimensional lasing system, consisting of a periodic metallic sheet supporting plasmonic modes, coated with a lasing material. The dynamics of the lasing phenomenon is studied as a function of the size of the pump laser which generates the population inversion in the lasing material. The Authors find that the response time of the laser, i.e. the delay between the input pump pulse and the output laser pulse, decreases with the size of the pump beam.*

*I find that this manuscript is very well written, with a concise and instructive introductory section. The main result of this work, i.e. the dependence of the response time on the pump beam size, is clearly stated. The Authors also motivate why the study of the response time is important for optoelectronic devices. I recommend this paper for publication, but I kindly ask the Authors to address the two points that I raise below.*

**Review**

*My main criticism of the paper concerns the statement that the calculations reveal an "optimal" beam size, yielding the shortest response time. From Fig.2(b) one sees that the shortest response time corresponds to a very flat minimum in the red curve. I doubt the robustness of this minimum as the system's parameters, or the level of approximation of the theory, are changed. I strongly suggest that the Authors rephrase the text, claiming that d\~15um corresponds to the minimum size to achieve a short response time, and not the optimal size.*

**Response**

**1.** We rephrase the text claiming that $d \sim 15 \mu \text{m}​$ corresponds to the minimum size to achieve the shortest response time, rather than the optimal size.

**Review**

*My second criticism is that the parameters used to calculate the results in Fig.2 are not given in the text. It is written that Fig.1 corresponds to the solution of Eqs.(1)-(3), but what is the size of the device, how many modes are considered, i.e. what are the values of the Rabi couplings, etc.? (The parameters are given only for the single-mode approximation.)*

**Response**

**2.** We have indicated the parameters of the plasmonic DFB laser in the main part of the manuscript. We have added a description of the mode structure of the plasmonic laser and a method for calculating the values of the constants (Rabi constant, the relaxation rates, etc.) from the laser equations (1)--(3).

### 2019-04-03 Response_DFB_Pulse_PRApplied.md
Modified by AZ and EA

@@
 *This paper theoretically studied the ultrafast response time of a plasmonic distributed feedback laser in the large-signal modulation regime.
-The authors demonstrated the dependence of response limit with*
-*pump spot size using both numerical and analytical approaches.
+The authors demonstrated the dependence of response limit with pump spot size using both numerical and analytical approaches.
 The identification on different ultrafast stages and insights on pump-tunable lasing dynamics could facilitate the applications of plasmon lasers in optoelectronic devices.*
@@
 *However, the manuscript focused on plasmonic nanohole arrays with a specific parameter design (100-nm hole, 470-nm spacing and 100-nm film thickness).
-Are the conclusions on the critical pump spot size and*
-*ultrafast response limit (1 ps) general enough to plasmon nanohole array lasers in different design parameters (film thickness, hole size, spacing)?*
+Are the conclusions on the critical pump spot size and ultrafast response limit (1 ps) general enough to plasmon nanohole array lasers in different design parameters (film thickness, hole size, spacing)?
+Also, plasmon lasers based on nanohole arrays and nanoparticle arrays show different lasing mechanisms from surface plasmon polariton and surface lattice resonances, respectively.
+It is not a fair comparison between the calculated response time of plasmon nanohole arrays and experimental results of nanoparticle arrays to verify the accuracy of theoretical modeling (final paragraph, the introduction section).*
 
@@
 
-**1.** The time response of the plasmonic laser is determined by the Purcell factor and the EM field relaxation rate in plasmonic structure.
-These characteristics depend on film thickness, hole size, spacing.
-For
-this reason, the values of the critical pump spot size and ultrafast response limit are different for divergent realizations of plasmon nanohole array lasers.
+Note that the pump spot effectively excites the EM field inside the pumped area which forms an effective laser mode.
+The existence of the *critical* size of the pump beam corresponds to the size of the effective mode.
+Qualitatively, the size of the effective mode is equal to the characteristic propagation length of the EM field along the plasmonic structure, which is$\sim v_{\text{group}}/ \gamma_a$.
+When the size of the pump beam is greater than the size of the effective laser mode, the response time almost does not depend on the size of the pump beam.
+When the size of the pump beam becomes smaller than the size of the effective laser mode, the response time increases with the decrease of the pump beam size.
 
-However, our analytical model shows that the existence of the critical pump spot size is related to finite size of an effective laser mode.
-When the size of pump beam is greater than the size of the effective
-laser mode the response time almost does not depend on the size of pump beam.
-When the size of the pump beam becomes smaller than the size of the effective laser mode, the response time increases with the decrease of the pump beam size.
-For this reason, the critical pump spot size takes place in different realizations of the plasmonic distributed-feedback lasers.
+The effective laser mode is formed by the Bloch modes of a periodic plasmonic structure which determine a characteristic group velocity and a dissipation constant.
+Of course, modes in different types of structure have different group velocities and relaxation constants.
+Nevertheless, from the qualitative arguments presented above, there is always the *critical* size of the pump beam, for which the response time is minimal.
 
-**Review**
+However, the specific value of the time response strongly depends on the type of the plasmonic structure.
+Indeed, the response time depends on the Rabi constant of interaction.
+In turn, the Rabi constant is determined by the volume of the plasmonic mode which is different for surface plasmon-polariton and surface lattice resonances.
+In our numerical simulations, we consider plasmon lasers based on nanohole arrays, though in the work [[Daskalakis, Konstantinos S., et al.
+*Nano letters* 18.4 (2018): 2658-2665.](https://pubs.acs.org/doi/abs/10.1021/acs.nanolett.8b00531)] the experiment has been conducted for plasmon lasers based on nanoparticle arrays in the different geometry.
+In this regard, we agree with the referee that the comparison of the response times obtained in the experiment and in our simulations is not a fair verification of the accuracy of our theoretical modeling.
 
-*Also, plasmon lasers based on nanohole arrays and nanoparticle arrays show different lasing mechanisms from surface plasmon polariton and surface lattice resonances, respectively.
-It is not a fair comparison*
-*between the calculated response time of plasmon nanohole arrays and experimental results of nanoparticle arrays to verify the accuracy of theoretical modeling (final paragraph, the introduction section).*
-
-**Response**
-
-**2.** It is true that in our numerical simulations we consider plasmon lasers based on nanohole arrays, though in the work [Torma] the experiment has been conducted for plasmon lasers based on nanoparticle arrays in the different geometry.
-In this regard, the conformity of the response times obtained in the experiment and in our simulations is not a fair verification of the accuracy of our theoretical modeling.
-We agree with the referee in this point and have reformulated the last paragraph in the Introduction section by excluding this direct comparison.
+In the revised version of the manuscript, we reformulate the last paragraph in the <u>Introduction</u> section by excluding this direct comparison between our theory and the experiment.
+In <u>Conclusion</u>, we add the qualitative arguments that the *critical* size of the pump beam is a common feature for plasmonic DFB lasers.
 
@@
 
-**3.** In our modeling we take into account a large amount of atoms, which have different coordinates $\textbf{r}_m$.
-The coupling constant between $j$th mode of the plasmonic structure and $m$th atom depends on atom's coordinates as
+In our modeling, we take into account a large number of atoms, which have different coordinates $\textbf{r}_m​$.
+The coupling constant between $j​$th mode of the plasmonic structure and $m​$th atom depends on atom's coordinates as
 $$
-\Omega_{jm} = \Omega_0 e^{i \textbf{k}_j \textbf{r}_m}
+\Omega _{jm} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar
 \tag{1}
@@
 
-The size of the pump beam and the power density of pumping are incorporated in the model as the initial conditions for population inversions of atoms.
-The initial value of the population inversion for the atoms located inside the pumping spot is $D_0$.
-Other atoms have initial population inversions equal to $-1$, i.e.
+where the electric field is taken at the position of the $m$th atom. The Rabi constants enter as coefficients in the Eqs. (1) – (3).
+
+The pulse pump of the active medium enters as an initial condition for the atomic population inversion.
+The initial value of the population inversion for the atoms whose coordinates lie inside the area of the pumping spot is $D_0$.
+Initial population inversions of other atoms equal to $-1$, i.e.
 they are in the ground state.
-The value of $D_0$ is obtained from the pumping power density as
+The *size* of the pump beam enters as an area, inside which atoms are excited.
+
+Pump power density (excitation fluence) enters through the value of $D_0​$. The connection between them has the form 
 $$
-\frac{D_0 + 1}{2} = \frac{P_0 \Delta t}{n_c S l \hbar \omega}
+\frac{D_0 + 1}{2} = \eta \frac{P_0 \Delta t}{n_c S l \hbar \omega}
 \tag{2}
@@
 Thus, we assume that the entire pulse energy comes to the excitation of atoms of the active medium.
-In real experiment, there are additional channels of energy losses, which lead to the decrease of the population inversion of the excited state.
-These losses can be taken into account as an addition factor in (2).
-For the simplicity, we do not consider the additional losses and use the estimation (2).
+In real experiment, there are additional circumstances as a result of which a part of the energy from pulse pumping does not come into the active medium.
+These circumstances are taken into account as an additional coefficient $\eta <1$ in a right-hand side of (2).
 
-In our simulations the power of the pump pulse is $P_0 \approx $. This value is typical for experiments carried out in with the same plasmonic structure in the same geometry, see [van exter, Beijnum].
+In our simulations **the power of the pump pulse** is $P_0 \approx ​$.
+This value is typical for experiments carried out with the same plasmonic structure in the same geometry, see [[F.
+Beijnum, P.
+Veldhoven, E.
+Geluk, M.
+Dood, G.
+Hooft, and M.
+van Exter, *Phys.
+Rev.
+Lett.* 110, 206802 (2013).](https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.110.206802)].
 
-Note that the dependence of the response time on the pump beam size is observable in a wide range of $D_0$ values, and a fine tune of the pumping parameters is not necessary to observe the predicted dependency.
+Note that the dependence of the response time on the pump beam size is observable in a wide range of $D_0$ values, and a fine tune of the pumping power is not necessary to observe the predicted dependency.
+
+In the revised manuscript, we add explanations on how experimental conditions, such as pump size and pump power density  are incorporated in the modeling.
 
@@
 
-**4.** We take into account diffractive coupling between nanoholes in laser equations.
-For this purpose, firstly, we find the EM field distribution of the Bloch modes of the plasmonic structure within the coupled-mode theory.
-Using this framework we allow for the scattering of the surface plasmon-polariton waves with different Bloch wavenumbers on the nanoholes.
-We use the scattering rates found in Ref.
-\[V.
+Diffractive coupling between nanoholes is taken into account in laser equations (1) – (3).
+
+First, we have found the EM field distribution of the Bloch modes of the plasmonic structure within the coupled-mode theory.
+Note that the first order of the coupled-mode theory describes the hybridization of initially non-interacting SPP modes through scattering on the holes, i.e.
+takes into account diffractive coupling **[]**.
+Scattering rates have been taken from Ref.
+[[V.
 T.
@@
 P.
-van Exter, *Journal of Optics* 16, 114019 (2014)\] for the plasmonic laser based on Au film perforated by the nanohole arrays.
-The coupled-mode theory with such scattering rates demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser \[V.
+van Exter, *Journal of Optics* 16, 114019 (2014)](https://iopscience.iop.org/article/10.1088/2040-8978/16/11/114019/meta)] for the plasmonic laser based on Au film perforated by the nanoholes array.
+The coupled-mode theory with such scattering rates demonstrates good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser [[V.
 T.
@@
 P.
-van Exter, Journal of Optics 16, 114019 (2014)\].
+van Exter, *Journal of Optics* 16, 114019 (2014)](https://iopscience.iop.org/article/10.1088/2040-8978/16/11/114019/meta)].
 
-Secondly, we use the obtained EM field distributions to calculate the Rabi constants between the Bloch modes and atoms of the active medium.
-The eigenfrequencies and the relaxation rates of the Bloch modes enter in the laser equations (1)-(3).
-
-Thus, the parameters and dispersive dependences of the Bloch modes of the plasmonic structure taken into account in the laser equations via the Rabi constants, the eiegnfrequencies, and the relaxation rates.
-
-We added the description of the band-structure of the plasmonic nanoholes array laser in the Section 2 of main part of the manuscript and **<u>Appendixes A and B.</u>**
-
-A change in the pump beam size results in a change in the volume of the collective mode.
-As a result, the Purcell factor and the rates of the spontaneous and stimulated emissions also change.
+Second, we use the obtained EM field distributions to calculate the Rabi constants between the Bloch modes and atoms of the active medium.
+These Rabi constants, the eigenfrequencies and the relaxation rates of the Bloch modes enter the laser equations (1) – (3).
+The Eqs.
+(1) – (3) determine the amplitude of each Bloch mode.
+The amplitudes depend on the interaction between the Bloch mode and active atoms, and, as a consequence, on the pump beam size.
+The field of the collective mode is formed by the large number of Bloch modes of a periodic plasmonic structure.
+As a result, the volume of the collective mode changes with a change in the pump beam size.
+The Purcell factor and the rates of the spontaneous and stimulated emissions also change.
 The strongest change occurs when the diameter of the pump beam is less than the critical value, at which the strongest dependence of the response time on the diameter appears.
@@
 Note that in our modeling the diameter of the pump beam is always much larger than the period of the nanohole array.
-For this reason, we do not take into account the inhomogeneous pumping within the cell of plasmonic structure.
+For this reason, we do not take into account effects on the boundary of the pump beam.
+
+In the revised manuscript, we add the description of the band-structure of the plasmonic nanoholes array laser in the <u>Section 2</u> of the main part of the manuscript and the <u>Appendix</u>.
 
@@
 
-**5.** We added the derivation of the Rabi constant through the parameters of the plasmonic structure and the active medium in the revised version of the manuscript , see Section 2 in main part of the text and the Appendix A.
+**Response**
+
+We added the derivation of the Rabi constant through the parameters of the plasmonic structure and the active medium in the revised version of the manuscript , see <u>Section 2</u> in main part of the text and the <u>Appendix</u>.
 
@@
 
-In our simulations we take into account several hundred of the Bloch modes of the plasmonic structure which form together the EM field distribution in the active medium.
-However, our simulations show that the resulting the EM field distribution slightly changes in the first and second stage of the output pulse evolution.
-In other words, there is a dynamical equilibrium between relaxation and propagation of the EM field on the one hand and the stimulated amplification of the EM field in the active medium on the other hand.
+In our simulations we take into account $\sim 500$ Bloch modes of the plasmonic structure which form together the EM field distribution in the active medium.
+However, our simulations show that the *resulting* EM field distribution slightly changes in the first and second stage of the output pulse evolution.
+In other words, there is a dynamical equilibrium between relaxation and propagation of the EM field outside the pump beam on the one hand, and the stimulated amplification of the EM field in the active medium on the other hand.
 We use the fact of the constancy of the EM field distribution to construct the single-mode model.
@@
 
-The change of the EM field distribution from Fig 3d to 3f is associated with a stronger decrease of the population inversion in the third stage of the response.
+The change of the EM field distribution from Fig.
+3d to 3f is associated with a stronger decrease of the population inversion in the third stage of the response.
 This leads to a violation of  the dynamical equilibrium between relaxation, propagation of the EM field and the stimulated amplification of the EM field in the system.
@@
 5b.
-We added scale bar to the Figure 3; we indicated in the figure caption the time
-moments corresponding to the intensity field distribution in Figures 3. We described the meaning of dashed circles in the Figure 3.
-
-We also changed the caption to Figure 3.
-
-In the previous version:
-
-"Field distribution of the collective mode when ..."
-
-In a new version:
-
-"Distribution of the EM field intensity on the surface of plasmonic DFB
-laser when ..."
+We added a scale bar to the Fig.
+3; we indicated in the figure caption the time
+moments corresponding to the intensity field distributions in Figs. 3. We described the meaning of dashed circles in the Fig. 3.
 
@@
 
-We placed all mentions to the superradiance problem to the Conclusion section of the manuscript and noted the potential interest in studying the similarity of the observed phenomenon and superradiance.
+We placed all mentions to the superradiance problem to the <u>Conclusion</u> section of the manuscript and noted the potential interest in studying the similarity of the observed phenomenon and superradiance.
 
@@
 I doubt the robustness of this minimum as the system's parameters, or the level of approximation of the theory, are changed.
-I strongly suggest that the Authors rephrase the text, claiming that d\~15um corresponds to the minimum size to achieve a short response time, and not the optimal size.*
+I strongly suggest that the Authors rephrase the text, claiming that $d \sim 15 \mu\text{m}​$ corresponds to the minimum size to achieve a short response time, and not the optimal size.*
 
@@
 
-**1.** We rephrase the text claiming that $d \sim 15 \mu \text{m}​$ corresponds to the minimum size to achieve the shortest response time, rather than the optimal size.
+We agree with the referee that the minimum is flat and some unaccounted effects, such as non-ideal plasmonic structure, inhomogeneous distribution of the active medium, etc., may change the value of the size beam for which the response time is minimal.
+We rephrase the text claiming that $d \sim 15 \mu \text{m}​$ corresponds to the minimum size to achieve the short response time for considered structure, rather than the optimal size.
 
@@
 **2.** We have indicated the parameters of the plasmonic DFB laser in the main part of the manuscript.
-We have added a description of the mode structure of the plasmonic laser and a method for calculating the values of the constants (Rabi constant, the relaxation rates, etc.) from the laser equations (1)--(3).
+We have added a description of the mode structure of the plasmonic laser and a method for calculating the values of the constants (Rabi constant, the relaxation rates, etc.) from the laser equations (1) – (3).
+We also add <u>Appendix</u> in the revised manuscript.

### 2019-04-03 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Modified by AZ and EA

@@
 In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
-Consequently, there is a critical minimum diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the shortest response time.
+Consequently, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
 This diameter is about the decay length of the EM field in the plasmonic structure and for the plasmonic laser under consideration is $\sim 15 \mu$m.
@@
 \begin{equation}\label{eq4SM}
-{\Omega _{jm}} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar,
+{\Omega _{jm}} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar.
 \end{equation}
-where ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom.
-The electric field `per one quantum' is inversely proportional to the mode volume (see Appendix A).
-As a result, the Rabi constant increases with the decrease of the mode volume.
-The Eqs.~(\ref{eq1})--(\ref{eq3}) allow for the dependence of the stimulated and spontaneous emissions rates on the Rabi constant that enables to take into account the Purcell effect of the plasmonic structure.
-Every atom of the active medium has its coordinates, which determine the Rabi constants between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
+Here ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom.
+Every atom of the active medium has its coordinates, which determine the Rabi constants of interaction between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
 This allows us to take into account the spatial distribution of the pumping and the population inversion within the active medium.
+Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition, $D_m(0)$ for atoms lying inside the pump beam.
+The value of $D_m(0)$ is proportional to the pumping energy per unit area.
+The population inversions of the atoms whose coordinates lie outside the pump beam are $-1$ at the initial time.
+Note that the square of the Rabi constant is inversely proportional of the Bloch mode volume and is proportional of the Purcell factor of the mode (see Appendix).
 
@@
 The surface of the entire periodic plasmonic structure is $190 \times 190$ $\mu \text{m}^2$.
-We consider that the transition frequency of the active medium is close to the frequency of the second-order Bragg grating in the plasmonic structure and the linewidth of the active medium is about 100 nm.
+We consider that the transition frequency of the active medium, $\omega_\sigma$, is close to the frequency of the second Bloch condition in the plasmonic structure, i.e., $k_{pl}(\omega_{\sigma}) \approx G$, where $k_{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector.
+\cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} and the linewidth of the active medium is about 100 nm.
 
@@
 
-In the absence of the holes, the eigenmodes of the plasmonic structure are TM-modes, similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
-These modes have an in-plane component of the magnetic field, $H_{\parallel}$, which is perpendicular to the direction of mode's propagation and both out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}.
-The in-plane electric field, $E_{\parallel}$, is in the direction of the mode's propagation. 
-The out-of-plane component of the electric field, $E_{\perp}$, is much greater than the in-plane component, $E_{\parallel}$ \cite{TennerJOpt,TennerACSPhot}.
-We find the eigenmodes of the plasmonic structure with the holes within the coupled-mode theory with the experimental values of the scattering rates (see~\cite{TennerJOpt} and Section 2 in Suppl.Mat.
-of \cite{nefedkin2018acsphot}).
-Corresponding dispersion curves are shown in \textbf{\underline{in Fig.~.}} The modes corresponding to different dispersion curves have various radiation losses and field distributions inside the unit cell of the plasmonic crystal (see Section 2 in Suppl.Mat.
-of \cite{nefedkin2018acsphot}).
-The far-field radiation from the plasmonic structure arises mainly due to the scattering of $H_{\parallel}$ on the holes~\cite{TennerJOpt}.
-For this reason, the modes that have a minimum value of $H_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $H_{\parallel}$ at the holes.
-In the continuous-wave regime, the lasing takes place at the modes with small radiation losses and a high quality factor lying on the 2nd dispersion curve, \textbf{\underline{see Fig.~}} \cite{TennerJOpt,nefedkin2018acsphot}.
+The EM field distribution of the Bloch modes of the plasmonic structure have been found within the coupled-mode theory \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} (see also Appendix) which describes hybridization of modes of the system without holes through scattering on the holes.
+Scattering rates for these modes have been taken from \cite{TennerJOpt,TennerACSPhot}.
+The coupled-mode theory with such scattering rates demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole array laser \cite{TennerJOpt,TennerACSPhot}.
 
-In numerical simulations, we take into account several hundreds of the Bloch modes of the plasmonic structure lying on different dispersion curves.
+In numerical simulations, we take into account $468$ Bloch modes of the plasmonic structure lying on different dispersion curves.
 The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat.
 of \cite{nefedkin2018acsphot}.
-
-We consider the response of the system to ultrashort pulse pumping of the active medium.
-Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}).
-We consider that at an initial time the population inversions of the atoms whose coordinates lie inside the pump beam are expressed as
-\begin{equation}\label{eq5SM}
-{\left(D_m+1\right)/2} = {P_0 \Delta t / \left(n_c S h \hbar \omega \right)}. 
-\end{equation}
-Here $\left(D_m+1\right)/2$ is a population of the upper level of the atom; $P_0$ is a pump power; $\Delta t$ is a duration of the pump pulse; $n_c$ is a concentration of the active atoms; $S$ is an area of the pump beam; $h$ is the thickness of Au film; $\omega$ is a frequency of the pump pulse.
-Also, we consider that the population inversions of the atoms whose coordinates lie outside the pump beam are $-1$ at the initial time.
 
@@
 Although in our simulations we take into account several hundreds of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response.
-In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field and the stimulated amplification of the EM field in the active medium.
+In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field outside the pumped area and the stimulated amplification of the EM field inside the pumped area.
 When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
@@
 We use the fact of the constancy of the EM field distribution to construct a single-mode model.
-To do this, we assume that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multi-mode problem.
+To do this, we assume that there is a single collective mode which EM field distribution coincides with the EM field distribution found in the numerical simulations of the multi-mode problem.
 Furthermore, we assume that the pumped active medium interacts only with such a single collective mode.
@@
 In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.
+It is interesting to note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
+Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).
+In the plasmonic cavity, the inequality $\gamma_a \gg \gamma_D$ is always hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds.
+We believe that further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.
 
-Thus, there is a critical minimum diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the shortest response time.
+Thus, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
 This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
@@
 Thus, the obtained results pave the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.
-
-It is interesting to note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
-Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).
-In the plasmonic cavity, the inequality $\gamma_a \gg \gamma_D$ is always hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds.
-We believe that further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.
 
@@
 where $A_{0j}$ is a real normalization constant for \textit{j}th mode; ${{\bf{e}}_j}\left( {x,y,z} \right) = {{\bf{e}}_j}\left( {x + {L_x},y + {L_y},z} \right)$  and ${{\bf{h}}_j}\left( {x,y,z} \right) = {{\bf{h}}_j}\left( {x + {L_x},y + {L_y},z} \right)$, where $L_x$ and $L_y$ are periods of the plasmonic structure along $x$-axis and $y$-axis, respectively.
+
+The functions ${{\bf{e}}_j}\left(\textbf{r}\right)$ and ${{\bf{h}}_j}\left(\textbf{r}\right)$ can be find within the couple-mode theory, see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat of \cite{nefedkin2018acsphot}.
+In this approach the eigenmodes of the plasmonic structure with holes are expressed as a linear combinations of the eigenmodes of the plasmonic structure without holes \cite{TennerJOpt,TennerACSPhot}.
+The expansion coefficients are calculated through the scattering matrix which contains the scattering amplitudes at angles $0^{\circ }$, $\pm 90^{\circ }$ and $180^{\circ }$ and effective refractive index of the plasmonic structure \cite{TennerJOpt,TennerACSPhot}.
+
+The eigenmodes of the plasmonic structure without holes are TM-modes, similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
+These modes have an in-plane component of the magnetic field, $h_{\parallel}$, which is perpendicular to the direction of mode's propagation and both out-of-plane, $e_{\perp}$, and in-plane, $e_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}.
+The in-plane electric field, $e_{\parallel}$, is in the direction of the mode's propagation. 
+The out-of-plane component of the electric field, $e_{\perp}$, is much greater than the in-plane component, $e_{\parallel}$ \cite{TennerJOpt,TennerACSPhot} and so it plays main role at the interaction of the modes with the active medium.
+
+The eigenmodes of the plasmonic structure with the holes we find within the coupled-mode theory with the experimental values of the scattering amplitudes, see~\cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat.
+of \cite{nefedkin2018acsphot}.
+The coupled-mode theory with such scattering amplitudes demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser \cite{TennerJOpt}.
+At close to the transition frequency of the active medium there are four different dispersion curves.
+The eigenmodes corresponding to different dispersion curves have various radiation losses and EM field distributions inside the unit cell of the plasmonic structure (see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat.
+of \cite{nefedkin2018acsphot}).
+In particularly, at the edge of the band gap when the Bloch wavevectors, $\left(k_{xj}, k_{yj}\right) = \left(0, 0\right)$  the EM field distributions in the modes from the different dispersion curves are
+
+\textbf{Table A}
+
+$\begin{array}{cccc}
+\text{The number of the dispersive curve} & e_{\parallel} & \
+h_{\parallel} & \
+e_ {\perp } \left( e_z \right) \\
+\text{1} & \sin (G x)+\sin (G y) & \sin (G x)+\sin (G y) & \
+\cos (G x)+\cos (G y) \\
+\text{2} & \sin (G x)-\sin (G y) & \sin (G x)-\sin (G y) & \
+\cos (G x)-\cos (G y) \\
+\text{3} & \cos (G y) & \cos (G y) & \sin (G y) \\
+\text{4} & \cos (G x) & \cos (G x) & \sin (G x) \\
+\end{array}$
+\\
+see~\cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}.
+
+The far-field radiation from the plasmonic structure arises mainly due to the scattering of $h_{\parallel}$ on the holes~\cite{TennerJOpt}. 
+For this reason, the modes that have a minimum value of $h_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $h_{\parallel}$ at the holes.
+Note that in the continuous-wave regime, the lasing takes place at the modes with small radiation losses and a high quality factor lying on the 2nd dispersion curve, \cite{TennerJOpt,nefedkin2018acsphot}.
+\\
 
@@
 Also, ${f_j}\left( {{{\bf{r}}_m}} \right) = \left( {{{\bf{d}}_m} \cdot {{\bf{e}}_j}\left( {{{\bf{r}}_m}} \right)} \right)/(\left|\bf{d}_m\right| \left|\textbf{e}_{j}\right|_{max})$ is a periodic function with plasmonic lattice cell period.
-Note that the functions ${f_j}\left( {{{\bf{r}}_m}} \right)$ for the modes lying in different bands are orthogonal to each other.
-The orthogonality arises due to difference of the EM field distributions within the cell of the plasmonic structure for the modes lying in different bands.
-
-\section*{Appendix B}
-The eigenimodes of such a plasmonic structure without holes are the TM-modes that are similar to surface plasmons at the metal-dielectric interface.
-These modes have an in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of the mode propagation, and out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field.
-The in-plane electric field, $E_{\parallel}$, is in the direction of the mode propagation.
-For the surface plasmons the out-of-plane component of the electric field, $E_{\perp}$, is much greater than the in-plane component, $E_{\parallel}$.
-
-In a periodic lattice, the electric and magnetic fields of the eigenmodes satisfy the Bloch condition. 
-We consider that the transition frequency of the active medium is close to the frequency at which the second Bloch condition is satisfied, i.e.
-$k_{sp}(\omega_{\sigma}) \approx G$, where $k_{sp}$ is a wavevector of the TM-mode of the plasmonic structure without holes and $G$ is a reciprocal lattice vector.
-The eigenmodes of the 2D plasmonic DFB laser can be calculated by the coupled-mode theory~\cite{ExterOptExp,TennerJOpt}. 
-Within this framework the electric field is expressed as
-%\begin{widetext}
-\begin{equation}
-\begin{array}{l}
-{\bf{E}}\left( {r,t} \right) = \exp \left( {i{k_x}x + i{k_y}y} \right) \times \\
-\times \left( {{E_x}\left( t \right){{\bf{u}}_x}\exp \left( {iGx} \right) + {E_{ - x}}\left( t \right){{\bf{u}}_{ - x}}\exp \left( { - iGx} \right) + {E_y}\left( t \right){{\bf{u}}_y}\exp \left( {iGy} \right) + {E_{ - y}}\left( t \right){{\bf{u}}_{ - y}}\exp \left( { - iGy} \right)} \right)\tag{5}
-\end{array}
-\end{equation}
-%\end{widetext}
-Here, we suppose that the surface of the Au film lies perpendicularly to $z$ axis and the sides of the plasmonic lattice are in the direction of $x$ and $y$ axes. 
-The centre of the unit cell coincides with the centre of a hole.
-The vectors ${{\bf{u}}_x}$,  ${{\bf{u}}_{ - x}}$, ${{\bf{u}}_y}$, and ${{\bf{u}}_{ - y}}$  are the eigenvectors of propagating TM-modes of the plasmonic system without holes with respectively directed wavevectors.
-${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$  are the amplitudes of the corresponding eigenmodes, and ${k_{x,y}} = \omega /c\sin {\theta _{x,y}}$, where ${\theta _{x,y}}$ is the angle of emission along the respective axes.
-
-The amplitudes  ${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$ are found from
-\begin{equation}
-dE/dt = i\hat WE = i\left( {{{\hat W}_0} + \hat V} \right)E\tag{6}
-\end{equation}
-where 
-%\begin{strip}
-\begin{equation}
-\begin{array}{l}
-{\hat W_0} = \left( {\begin{array}{*{20}{c}}
-	{{\omega _0}}&\gamma &\kappa &\kappa \\
-	\gamma &{{\omega _0}}&\kappa &\kappa \\
-	\kappa &\kappa &{{\omega _0}}&\gamma \\
-	\kappa &\kappa &\gamma &{{\omega _0}}
-	\end{array}} \right) , \\
-\hat V\left( {{\theta _x},{\theta _y}} \right) = \left( {\begin{array}{*{20}{c}}
-	{{c_2}\theta _x^2 + {c_1}{\theta _y}}&0&0&0\\
-	0&{{c_2}\theta _x^2 - {c_1}{\theta _y}}&0&0\\
-	0&0&{{c_1}{\theta _x} + {c_2}\theta _y^2}&0\\
-	0&0&0&{ - {c_1}{\theta _x} + {c_2}\theta _y^2}
-	\end{array}} \right)\tag{7}
-\end{array}
-\end{equation}
-%\end{strip}
-Here, ${\omega _0} = Gc/{n_\text{eff}}$ ($c$ is the speed of light), $\gamma $  and $\kappa$  are the back-scattering amplitude at angle $180^{\circ }$ and the right-angle scattering at $\pm 90^{\circ }$, respectively; ${c_1} = {\omega _0}/{n_\text{eff}}$, ${c_2} = {\omega _0}/\left( {2n_\text{eff}^2} \right)$, and ${n_\text{eff}}$  is an effective refractive index~\cite{ExterOptExp,TennerJOpt}.
-The eigenvectors of $\hat H$  estimate the amplitudes ${E_x}$ , ${E_{ - x}}$,  ${E_y}$, and ${E_{ - y}}$.
-The values of $\gamma $, $\kappa$, $c_1$, $c_2$ and $n_\text{eff}$ are taken from~\cite{TennerJOpt}.
-Using them, we obtain the dispersion curves for the 2D DFB laser, see \textbf{Figure 2}.
-
-At $\theta _x = \theta _y = 0$ the eigenvectors of $\hat W$ are \text{(1,1,1,1)}, \text{(1,1,-1,-1)}, \text{(0,0,1,-1)} and \text{(1,-1,0,0)}. 
-These eigenvectors correspond to the following electric field distributions:
-
-\textbf{Table S1.}
-
-$\begin{array}{cccc}
-\text{The eigenvector} & E_{\parallel} & \
-H_{\parallel} & \
-E_ {\perp } \left( E_z \right) \\
-\text{(1,1,1,1)} & \sin (G x)+\sin (G y) & \sin (G x)+\sin (G y) & \
-\cos (G x)+\cos (G y) \\
-\text{(1,1,-1,-1)} & \sin (G x)-\sin (G y) & \sin (G x)-\sin (G y) & \
-\cos (G x)-\cos (G y) \\
-\text{(0,0,1,-1)} & \cos (G y) & \cos (G y) & \sin (G y) \\
-\text{(1,-1,0,0)} & \cos (G x) & \cos (G x) & \sin (G x) \\
-\end{array}$
-
-The dispersion curves in the main text are numbered in the same order as in Table 1.
-The modes radiate the far-fields because of the scattering of electromagnetic waves by the holes~\cite{TennerJOpt}. 
-The TM-modes dominantly couple with far-field via the in-plane magnetic field in the holes.
-Therefore, the first and second modes which have minimum of the in-plane magnetic field in the centre of the unit cell of the plasmonic lattice (see Table 1) do not radiate, i.e. they are dark modes. 
-On the contrary, the third and fourth modes are bright modes and they radiate.
+Note that the functions ${f_j}\left( {{{\bf{r}}_m}} \right)$ for the modes lying in the different bands are orthogonal to each other.
+The orthogonality arises due to difference of the EM field distributions within the cell of the plasmonic structure for the modes lying in the different bands, see Table A.
 

### 2019-04-03 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Modified by NN, t_delay corrections

@@
 To overcome losses in the metal, an external pumping (optical or electrical) of the active medium is used.
-In the case of optical pumping, the diameter of the pump beam is usually smaller than the size of the periodic plasmonic structure \cite{TennerACSPhot,ZhouNatNano,daskalakis2018nanolett}, see Fig.~\ref{fig1}a.
+In the case of optical pumping, the diameter of the pump beam is usually smaller than the size of the periodic plasmonic structure \cite{TennerACSPhot,ZhouNatNano,daskalakis2018nanolett}, see Fig.~\ref{fig1}.
 In such a case, the EM field is excited inside the pumped area and is formed by a large number of Bloch modes \cite{TennerACSPhot,nefedkin2018acsphot}.
@@
 \end{equation}
-Here ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom.
+Here ${{\bf{d}}_m}$ is the dipole moment of $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in $j$th mode at the position of $m$th atom.
 Every atom of the active medium has its coordinates, which determine the Rabi constants of interaction between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
 This allows us to take into account the spatial distribution of the pumping and the population inversion within the active medium.
-Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition, $D_m(0)$ for atoms lying inside the pump beam.
+Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all the characteristic times of the plasmonic DFB laser, in our model, the ultrashort pulse pumping is described as the initial condition, $D_m(0)$ for atoms lying inside the pump beam.
 The value of $D_m(0)$ is proportional to the pumping energy per unit area.
@@
 The Eqs.~(\ref{eq1})--(\ref{eq3}) generalize the rate equations widely used in laser theory \cite{SiegmanLasers}.
-In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the atoms, and the interference between the EM fields of the different modes, $n_{jk}$ \cite{Zyablovsky2017approach}.
+In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the Bloch modes, and the interference between the EM fields of the different modes, $n_{jk}$ \cite{Zyablovsky2017approach}.
 
@@
 The surface of the entire periodic plasmonic structure is $190 \times 190$ $\mu \text{m}^2$.
-We consider that the transition frequency of the active medium, $\omega_\sigma$, is close to the frequency of the second Bloch condition in the plasmonic structure, i.e., $k_{pl}(\omega_{\sigma}) \approx G$, where $k_{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector.
-\cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} and the linewidth of the active medium is about 100 nm.
+We consider that the transition frequency of the active medium, $\omega_{\sigma}$, is close to the frequency of the second Bloch condition in the plasmonic structure, i.e.
+$k_\text{pl}(\omega_{\sigma}) \approx G$, where $k_\text{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot}, and the linewidth of the active medium is about $100 \text{nm}$.
 
@@
 
-The EM field distribution of the Bloch modes of the plasmonic structure have been found within the coupled-mode theory \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} (see also Appendix) which describes hybridization of modes of the system without holes through scattering on the holes.
+The EM field distribution of the Bloch modes of the plasmonic structure have been found within the coupled-mode theory \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} (see also Appendix), which describes the hybridization of modes of the system without holes through scattering on the holes.
 Scattering rates for these modes have been taken from \cite{TennerJOpt,TennerACSPhot}.
@@
 The photon number in the plasmonic DFB laser is small.
-For the system under consideration, the duration of this stage, i.e., the delay time, turns out to be less than a picosecond.
+For the system under consideration, the duration of this stage, i.e. the delay time, turns out to be less than a picosecond.
 In the second stage, the photon number in the plasmonic laser rapidly increases due to stimulated emission in the active medium.
@@
 However, the response times manifest a non-monotonic dependence on the size of the pump beam, see Fig.~\ref{fig2}b.
-It can be seen that there is an optimal size of the pump beam at which the response time is a minimum.
-For the considered plasmonic DFB laser the optimal diameter of the pump beam is about 15 $\mu$m.
+It can be seen that there is a critical size of the pump beam at which the response time is a minimum.
+For the considered plasmonic DFB laser, the critical diameter of the pump beam is about 15 $\mu$m.
 As the pump beam diameter decreases, the response time increases rapidly.
@@
 This kind of behavior of the plasmonic DFB laser takes place at different initial values of the population inversion of the active atoms.
-Thus, there is an optimal diameter of the pump beam, $d_{\text{beam}}^*$, at which an ultrafast light modulation in the plasmonic DFB laser is achieved.
+Thus, there is a diameter of the pump beam, $d_{\text{beam}}^*$, at which an ultrafast light modulation in the plasmonic DFB laser is achieved.
 
@@
 
-The existence of an optimal size of the pump beam can be explained with the example of a single-mode model.
+The existence of a critical size of the pump beam can be explained with the example of a single-mode model.
 To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
-Although in our simulations we take into account several hundreds of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response.
+Although in our simulations we take into account several hundred of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response.
 In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field outside the pumped area and the stimulated amplification of the EM field inside the pumped area.
@@
 We use the fact of the constancy of the EM field distribution to construct a single-mode model.
-To do this, we assume that there is a single collective mode which EM field distribution coincides with the EM field distribution found in the numerical simulations of the multi-mode problem.
+To do this, we assume that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multi-mode problem.
 Furthermore, we assume that the pumped active medium interacts only with such a single collective mode.
@@
 \begin{gather} 
-\dot n =  - {\gamma _a}n + 2{N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
+\dot n =  - 2 {\gamma _a}n + {N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
 \\
-\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
+\dot D =  - {\gamma _D}\left( {D + 1} \right) - 2\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
 \end{gather}
@@
 \begin{equation} 
-n\left( t \right) = \frac{{2{N_{\text{at}}}\Omega _R^2\left( {D + 1} \right)/{\gamma _\sigma }}}{{4{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma } - {\gamma _a}}}\left( {\exp \left( {4{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma } - {\gamma _a}} \right)t - 1} \right) = \frac{{\xi \left( {D + 1} \right)}}{{2\left( {\xi-1} \right)}}\left( {\exp \left( {\xi D - 1} \right){\gamma _a}t - 1} \right), \label{ph}
+n\left( t \right) = \frac{{{N_{\text{at}}}\Omega _R^2\left( {D + 1} \right)/{\gamma _\sigma }}}{{2{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma } - 2 {\gamma _a}}}\left( {\exp \left[ 2\left( {{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma } - {\gamma _a}} \right)t \right] - 1} \right) = \frac{{\xi \left( {D + 1} \right)}}{{2\left( {\xi-1} \right)}}\left( {\exp \left[\left( {\xi D - 1} \right) 2 {\gamma _a}t \right] - 1} \right), \label{ph}
 \end{equation}
-where $\xi={4{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma \gamma_a }}$ is the inverse threshold value of the pumping in the CW regime \cite{ScullyQO}.
+where $\xi={{N_{\text{at}}}\Omega _R^2D/{\gamma _\sigma \gamma_a }}$ is the inverse threshold value of the pumping in the continuous-wave (CW) regime \cite{ScullyQO}.
 At the initial moment, $D \simeq 1$ and we have  
 \begin{equation} 
-n\left( t \right) = \frac{\xi }{{\left( {\xi - 1} \right)}}\left( {\exp \left( {\xi  - 1} \right){\gamma _a}t - 1} \right). \label{ph0}
+n\left( t \right) = \frac{\xi }{{\left( {\xi - 1} \right)}}\left( {\exp \left[ \left( {\xi  - 1} \right) 2 {\gamma _a}t \right] - 1} \right). \label{ph0}
 \end{equation}
@@
 \begin{equation} 
-\dot D =  - {\gamma _D}\left( {1 + \eta } \right)\left( {D + 1} \right) - {\gamma _D}\frac{{2\xi \eta D}}{{\left( {\xi  - 1} \right)}}\left( {\exp \left( {\xi  - 1} \right){\gamma _a}t - 1} \right), \label{inv0}
+\dot D =  - {\gamma _D}\left( {1 + \eta } \right)\left( {D + 1} \right) - {\gamma _D}\frac{{2\xi \eta D}}{{\left( {\xi  - 1} \right)}}\left( {\exp \left[\left( {\xi  - 1} \right) 2 {\gamma _a}t \right] - 1} \right), \label{inv0}
 \end{equation}
-where we have put $\eta = 4 \Omega_R^2 /\gamma_D \gamma_\sigma$.
+where we have put $\eta = 2 \Omega_R^2 /\gamma_D \gamma_\sigma$.
 
@@
 \begin{equation} 
-\dot D =  - {\gamma _D}\frac{{2\xi \eta D}}{{\left( {\xi  - 1} \right)}}{\exp \left( {\xi  - 1} \right){\gamma _a}t} \label{invnl1}
+\dot D =  - {\gamma _D}\frac{{2\xi \eta D}}{{\left( {\xi  - 1} \right)}} \exp \left[\left( {\xi  - 1} \right) 2 {\gamma _a}t \right] \label{invnl1}
 \end{equation}
@@
 \begin{equation} 
-D\left( t \right) \simeq \exp \left( { - \frac{{2\xi \eta {\gamma _D}}}{{{{\left( {\xi  - 1} \right)}^2}{\gamma _a}}}\exp \left( {\xi  - 1} \right){\gamma _a}t} \right) = \exp \left( { - \frac{{2{\xi ^2}}}{{{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}}}\exp \left( {\xi  - 1} \right){\gamma _a}t} \right).
+D(t) \simeq \exp \left(  - \frac{{\xi \eta {\gamma _D}}}{{{\left( {\xi  - 1} \right)}^2}{\gamma _a}}  \exp \left[\left( {\xi  - 1} \right) 2 {\gamma _a}t \right] \right) = \exp \left(  - \frac{{{\xi ^2}}}{{{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}}}\exp \left[\left( {\xi  - 1} \right) 2 {\gamma _a}t \right] \right).
 \end{equation}
@@
 \begin{equation} 
-{t_{delay}} = \gamma _a^{ - 1}{\left( {\xi  - 1} \right)^{ - 1}}\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/2{\xi ^2}} \right). \label{tdele}
+{t_{delay}} = (2\gamma _a)^{ - 1}{\left( {\xi  - 1} \right)^{ - 1}}\ln \left( {{N_{\text{at}}}{{\left( {\xi - 1} \right)}^2}/{\xi ^2}} \right). \label{tdele}
 \end{equation}
 At this time, a sharp decrease of $D$ appears, see Fig.~\ref{fig4}b.
-Note that the assumption that the evolution of $D$ is much slower than the evolution of $n$ is valid when $t_{delay} \gg \gamma_a^{-1} \xi^{-1}$, i.e., $\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/2{\xi ^2}} \right) \gg 1$.
+Note that the assumption that the evolution of $D$ is much slower than the evolution of $n$ is valid when $t_{delay} \gg \gamma_a^{-1} \xi^{-1}$, i.e., $\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/{\xi ^2}} \right) \gg 1$.
 When $\xi \gg 1$, the assumption is correct if $N_{\text{at}} \gg 1$, which is true for the considered system.
@@
 \begin{equation} 
-{t_{\text{delay}}} = {\gamma _\sigma }\ln \left( {{N_{\text{at}}}/2} \right)/4{N_{\text{at}}}\Omega _R^2. \label{tdel}
+{t_{\text{delay}}} = {\gamma _\sigma }\ln \left( {{N_{\text{at}}}} \right)/2{N_{\text{at}}}\Omega _R^2. \label{tdel}
 \end{equation}
@@
 For the parameters of the DFB laser we have $d_{\text{beam}}^* \sim 10 \mu \text{m}$, which is in agreement with the full model, Fig.~\ref{fig2}b.
-Thus, because the delay time grows when $d_{\text{beam}}>d_{\text{beam}}^*$ and grows in the opposite case as well, we come to the conclusion that there is an optimal value, $\sim d_{\text{beam}}^*$, for the diameter of the pump beam for which the response time is a minimum.
+Thus, because the delay time grows when $d_{\text{beam}}>d_{\text{beam}}^*$ and grows in the opposite case as well, we come to the conclusion that there is a critical value, $\sim d_{\text{beam}}^*$, for the diameter of the pump beam for which the response time is a minimum.
 
@@
 The functions ${{\bf{e}}_j}\left(\textbf{r}\right)$ and ${{\bf{h}}_j}\left(\textbf{r}\right)$ can be find within the couple-mode theory, see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat of \cite{nefedkin2018acsphot}.
-In this approach the eigenmodes of the plasmonic structure with holes are expressed as a linear combinations of the eigenmodes of the plasmonic structure without holes \cite{TennerJOpt,TennerACSPhot}.
+In this approach, the eigenmodes of the plasmonic structure with holes are expressed as linear combinations of the eigenmodes of the plasmonic structure without holes \cite{TennerJOpt,TennerACSPhot}.
 The expansion coefficients are calculated through the scattering matrix which contains the scattering amplitudes at angles $0^{\circ }$, $\pm 90^{\circ }$ and $180^{\circ }$ and effective refractive index of the plasmonic structure \cite{TennerJOpt,TennerACSPhot}.
@@
 The coupled-mode theory with such scattering amplitudes demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser \cite{TennerJOpt}.
-At close to the transition frequency of the active medium there are four different dispersion curves.
+Near the transition frequency of the active medium, there are four different dispersion curves.
 The eigenmodes corresponding to different dispersion curves have various radiation losses and EM field distributions inside the unit cell of the plasmonic structure (see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat.
 of \cite{nefedkin2018acsphot}).
-In particularly, at the edge of the band gap when the Bloch wavevectors, $\left(k_{xj}, k_{yj}\right) = \left(0, 0\right)$  the EM field distributions in the modes from the different dispersion curves are
+In particular, at the edge of the band gap, when the Bloch wavevectors $\left(k_{xj}, k_{yj}\right) = \left(0, 0\right)$, the EM field distributions in the modes from the different dispersion curves are
 
@@
 For this reason, the modes that have a minimum value of $h_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $h_{\parallel}$ at the holes.
-Note that in the continuous-wave regime, the lasing takes place at the modes with small radiation losses and a high quality factor lying on the 2nd dispersion curve, \cite{TennerJOpt,nefedkin2018acsphot}.
+Note that in the CW regime, the lasing takes place at the modes with small radiation losses and high quality factors lying on the 2nd dispersion curve, \cite{TennerJOpt,nefedkin2018acsphot}.
 \\

### 2019-04-04 Response_DFB_Pulse_PRApplied.md
Final version of Response

@@
-----------------------------------------------------------------------
-
 # Report of Referee A -- NB10105/Nefedkin
-
-----------------------------------------------------------------------
 
@@
 Note that the pump spot effectively excites the EM field inside the pumped area which forms an effective laser mode.
-The existence of the *critical* size of the pump beam corresponds to the size of the effective mode.
-Qualitatively, the size of the effective mode is equal to the characteristic propagation length of the EM field along the plasmonic structure, which is$\sim v_{\text{group}}/ \gamma_a$.
+The existence of the *critical* size of the pump beam corresponds to the minimal size of the effective mode.
+Qualitatively, the minimal size of the effective mode is equal to the characteristic propagation length of the EM field along the plasmonic structure, which is$\sim v_{\text{group}}/ \gamma_a$.
 When the size of the pump beam is greater than the size of the effective laser mode, the response time almost does not depend on the size of the pump beam.
@@
 Of course, modes in different types of structure have different group velocities and relaxation constants.
-Nevertheless, from the qualitative arguments presented above, there is always the *critical* size of the pump beam, for which the response time is minimal.
+Nevertheless, based on the qualitative arguments presented above, we can conclude that there is always the *critical* size of the pump beam, for which the response time is minimal.
 
@@
 In our numerical simulations, we consider plasmon lasers based on nanohole arrays, though in the work [[Daskalakis, Konstantinos S., et al.
-*Nano letters* 18.4 (2018): 2658-2665.](https://pubs.acs.org/doi/abs/10.1021/acs.nanolett.8b00531)] the experiment has been conducted for plasmon lasers based on nanoparticle arrays in the different geometry.
+*Nano letters* 18.4 (2018): 2658-2665.](https://pubs.acs.org/doi/abs/10.1021/acs.nanolett.8b00531)] the experiment has been conducted for plasmon lasers based on nanoparticle arrays.
 In this regard, we agree with the referee that the comparison of the response times obtained in the experiment and in our simulations is not a fair verification of the accuracy of our theoretical modeling.
@@
 In the revised version of the manuscript, we reformulate the last paragraph in the <u>Introduction</u> section by excluding this direct comparison between our theory and the experiment.
-In <u>Conclusion</u>, we add the qualitative arguments that the *critical* size of the pump beam is a common feature for plasmonic DFB lasers.
+
+---
 
@@
 The pulse pump of the active medium enters as an initial condition for the atomic population inversion.
-The initial value of the population inversion for the atoms whose coordinates lie inside the area of the pumping spot is $D_0$.
-Initial population inversions of other atoms equal to $-1$, i.e.
+Because interaction of each atom with plasmon EM field depends on the atomic coordinate, a distribution of the initial population inversion over the atoms allows to determine the pump size.
+The initial value of the population inversion for the atoms whose coordinates lie inside the area of the pumping spot is $D_0​$.
+Initial population inversions of other atoms equal to $-1​$, i.e.
 they are in the ground state.
@@
 Here $P_0$ is power of the laser pulse pump; $\Delta t$ is duration of the pump pulse; $n_c$ is the concentration of atoms of the active medium; $S$ is the area of the pump beam; $l$ is the width of the active medium layer; $\hbar \omega$ is the transition frequency of the atom; and $(D_0 + 1)/2$ is the population of the excited state arising due to the pump pulse.
-
-Thus, we assume that the entire pulse energy comes to the excitation of atoms of the active medium.
 In real experiment, there are additional circumstances as a result of which a part of the energy from pulse pumping does not come into the active medium.
 These circumstances are taken into account as an additional coefficient $\eta <1$ in a right-hand side of (2).
-
-In our simulations **the power of the pump pulse** is $P_0 \approx ​$.
-This value is typical for experiments carried out with the same plasmonic structure in the same geometry, see [[F.
-Beijnum, P.
-Veldhoven, E.
-Geluk, M.
-Dood, G.
-Hooft, and M.
-van Exter, *Phys.
-Rev.
-Lett.* 110, 206802 (2013).](https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.110.206802)].
-
-Note that the dependence of the response time on the pump beam size is observable in a wide range of $D_0$ values, and a fine tune of the pumping power is not necessary to observe the predicted dependency.
+In our simulations, we assume that $P_0$ is such that all atoms inside the pumped area are inverted, namely we take the values from $D(0) = 0.25$ to $D(0) = 1$.
 
 In the revised manuscript, we add explanations on how experimental conditions, such as pump size and pump power density are incorporated in the modeling.
+
+---
 
@@
 First, we have found the EM field distribution of the Bloch modes of the plasmonic structure within the coupled-mode theory.
-Note that the first order of the coupled-mode theory describes the hybridization of initially non-interacting SPP modes through scattering on the holes, i.e.
-takes into account diffractive coupling **[]**.
+Note that the coupled-mode theory describes the hybridization of initially non-interacting SPP modes through scattering on the holes, i.e.
+takes into account diffractive coupling [[V.
+T.
+Tenner, A.
+N.
+van Delft, M.
+J.
+A.
+de Dood, and M.
+P.
+van Exter, *Journal of Optics* 16, 114019 (2014)](https://iopscience.iop.org/article/10.1088/2040-8978/16/11/114019/meta)].
 Scattering rates have been taken from Ref.
@@
 The Purcell factor and the rates of the spontaneous and stimulated emissions also change.
-The strongest change occurs when the diameter of the pump beam is less than the critical value, at which the strongest dependence of the response time on the diameter appears.
 
@@
 In the revised manuscript, we add the description of the band-structure of the plasmonic nanoholes array laser in the <u>Section 2</u> of the main part of the manuscript and the <u>Appendix</u>.
+
+---
 
@@
 
-We added the derivation of the Rabi constant through the parameters of the plasmonic structure and the active medium in the revised version of the manuscript , see <u>Section 2</u> in main part of the text and the <u>Appendix</u>.
+We added the derivation of the Rabi constant through the parameters of the plasmonic structure and the active medium in the revised version of the manuscript , see <u>Section 2</u> in main part of the text and <u>Appendix</u>.
+
+---
 
@@
 However, our simulations show that the *resulting* EM field distribution slightly changes in the first and second stage of the output pulse evolution.
-In other words, there is a dynamical equilibrium between relaxation and propagation of the EM field outside the pump beam on the one hand, and the stimulated amplification of the EM field in the active medium on the other hand.
+In other words, there is a dynamical equilibrium between relaxation and propagation of the EM field outside the pump beam, on the one hand, and the stimulated amplification of the EM field inside the pump beam, on the other hand.
 We use the fact of the constancy of the EM field distribution to construct the single-mode model.
@@
 
+---
+
 **2.** *Regarding the figures, the order of Figure 5b and 5a can be switched since 5b appeared earlier in the main text.
@@
 We added a scale bar to the Fig.
-3; we indicated in the figure caption the time
-moments corresponding to the intensity field distributions in Figs. 3. We described the meaning of dashed circles in the Fig. 3.
+3; we indicated in the figure caption the time moments corresponding to the intensity field distributions in Figs.
+3.
+We described the meaning of dashed circles in the Fig.
+3.
+
+---
 
@@
 
-----------------------------------------------------------------------
-
 # Report of Referee B -- NB10105/Nefedkin
-
-----------------------------------------------------------------------
 
@@
 We agree with the referee that the minimum is flat and some unaccounted effects, such as non-ideal plasmonic structure, inhomogeneous distribution of the active medium, etc., may change the value of the size beam for which the response time is minimal.
-We rephrase the text claiming that $d \sim 15 \mu \text{m}​$ corresponds to the minimum size to achieve the short response time for considered structure, rather than the optimal size.
+We rephrase the text claiming that $d \sim 15 \mu \text{m}$ corresponds to the minimum size to achieve the short response time for considered structure, rather than the optimal size.
+
+---
 

### 2019-04-04 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Final version of Response

@@
 	In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
-	We demonstrate that there is a critical minimum size of the pump beam for which the modulation speed is a maximum and can achieve $1$ THz.
+	We demonstrate that there is a minimum size of the pump beam for which the modulation speed is a maximum, and can achieve $1$ THz.
 	The obtained result opens the way to increasing both the modulation speed and the energy efficiency of plasmonic laser devices.
@@
 \end{equation}
-Here ${{\bf{d}}_m}$ is the dipole moment of $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in $j$th mode at the position of $m$th atom.
-Every atom of the active medium has its coordinates, which determine the Rabi constants of interaction between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
+Here ${{\bf{d}}_m}$ is the dipole moment of $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in $j$th Bloch mode at the position $\textbf{r}_m$ of $m$th atom, see Appendix and Section 2 in Suppl.Mat.
+of \cite{nefedkin2018acsphot}.
+Every atom of the active medium has its coordinates $\textbf{r}_m$, which determine the Rabi constants of interaction between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}).
 This allows us to take into account the spatial distribution of the pumping and the population inversion within the active medium.
-Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all the characteristic times of the plasmonic DFB laser, in our model, the ultrashort pulse pumping is described as the initial condition, $D_m(0)$ for atoms lying inside the pump beam.
-The value of $D_m(0)$ is proportional to the pumping energy per unit area.
+Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all the characteristic times of the plasmonic DFB laser, in our model, the ultrashort pulse pumping is described as the initial condition, $D_m(0)$.
+For atoms lying inside the pump beam, the value of $D_m(0)$ is proportional to the pumping energy per unit area.
 The population inversions of the atoms whose coordinates lie outside the pump beam are $-1$ at the initial time.
@@
 We consider that the transition frequency of the active medium, $\omega_{\sigma}$, is close to the frequency of the second Bloch condition in the plasmonic structure, i.e.
-$k_\text{pl}(\omega_{\sigma}) \approx G$, where $k_\text{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot}, and the linewidth of the active medium is about $100 \text{nm}$.
-
-A similar plasmonic DFB laser have been implemented in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
+$k_\text{pl}(\omega_{\sigma}) \approx G$, where $k_\text{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
+The linewidth of the active medium is about $100 \text{nm}$.
+Similar plasmonic DFB lasers have been implemented in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and their parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
 
 The EM field distribution of the Bloch modes of the plasmonic structure have been found within the coupled-mode theory \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} (see also Appendix), which describes the hybridization of modes of the system without holes through scattering on the holes.
-Scattering rates for these modes have been taken from \cite{TennerJOpt,TennerACSPhot}.
-The coupled-mode theory with such scattering rates demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole array laser \cite{TennerJOpt,TennerACSPhot}.
+Scattering amplitudes for these modes have been taken from \cite{TennerJOpt,TennerACSPhot}.
+The coupled-mode theory with such scattering amplitudes demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole array laser \cite{TennerJOpt,TennerACSPhot}.
 
-In numerical simulations, we take into account $468$ Bloch modes of the plasmonic structure lying on different dispersion curves.
-The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat.
+In numerical simulations, we take into account $468$ Bloch modes which have different distributions of the EM field inside the cell of the plasmonic structure.
+The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure, see \cite{nefedkin2018acsphot} and Suppl.Mat.
 of \cite{nefedkin2018acsphot}.
@@
 We use the fact of the constancy of the EM field distribution to construct a single-mode model.
-To do this, we assume that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multi-mode problem.
+To do this, we assume that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multimode problem.
 Furthermore, we assume that the pumped active medium interacts only with such a single collective mode.
@@
 	\end{subfigure}
-	\caption{(a) The dependence of the diameter of the collective mode on the pump beam diameter (blue line).
-The black dashed line denotes the pump beam diameter.
-(b) The dependence of the delay time on the pump beam diameter.
+	\caption{(a) The dependence of the delay time on the pump beam diameter.
 The blue solid line is obtained from numerical simulation of Eqs.
@@
 (\ref{tdel}).
+(b) The dependence of the diameter of the collective mode on the pump beam diameter (blue line).
+The black dashed line denotes the pump beam diameter.
 	}
@@
 
-The figure \ref{fig5}b shows that the analytical approximation, Eq.~(\ref{tdel}), agrees well with the numerical simulation of the exact model, Eqs. (\ref{eq1})--(\ref{eq3}).
+The figure \ref{fig5}a shows that the analytical approximation, Eq.~(\ref{tdel}), agrees well with the numerical simulation of the exact model, Eqs. (\ref{eq1})--(\ref{eq3}).
 Let us analyze the dependence of the delay time, Eq.~(\ref{tdel}), on the size of the pumped area.
@@
 So, we have $t_{\text{delay}} \sim d_{\text{mode}}^2\ln \left( d_{\text{beam}}^2 \right ) /d_{\text{beam}}^2$.
-The dependence of $d_{\text{mode}}$ on $d_{\text{beam}}$ is shown in Fig.~\ref{fig5}a.
+The dependence of $d_{\text{mode}}$ on $d_{\text{beam}}$ is shown in Fig.~\ref{fig5}b.
 
@@
 When $d_{\text{beam}}>d_{\text{beam}}^*$, the electric field of the collective mode is localized inside the pump beam, see Figs.~\ref{fig3}(d)--(f).
-As a consequence, the mode diameter is proportional to the beam diameter, i.e., $d_{\text{mode}} \sim d_{\text{beam}}$, cf. Fig.~\ref{fig5}a.
+As a consequence, the mode diameter is proportional to the beam diameter, i.e., $d_{\text{mode}} \sim d_{\text{beam}}$, see Fig.~\ref{fig5}b.
 For this case, from Eq.~(\ref{tdel}) we have $t_{\text{delay}} \sim \ln d_{\text{beam}}$.
@@
 When $d_{\text{beam}}<d_{\text{beam}}^*$, the electric field of the collective mode is localized outside the pump beam, see Figs.~\ref{fig3}(a)--(c), and the mode size is determined by the characteristic propagation length of the EM field in the unpumped region.
-As a result, the mode diameter does not depend on the beam diameter, and $d_{\text{mode}} \sim \text{const}$, see Fig.~\ref{fig5}a.
+As a result, the mode diameter does not depend on the beam diameter, and $d_{\text{mode}} \sim \text{const}$, see Fig.~\ref{fig5}b.
 In this case, $t_{\text{delay}} \sim \ln(d_{\text{beam}}^2)/d_{\text{beam}}^2$.
 This is in agreement with an increase of the delay time  when the pump beam diameter of the plasmonic DFB laser is reduced, see Fig.~\ref{fig2}b.
-Note that the critical value $d_{\text{beam}}^*$ corresponds with the characteristic propagation length of the EM field outside the pumped active medium and may be evaluated as $d_{\text{beam}}^* \sim v_{\text{group}}/\gamma_a$.
+Note that the critical value $d_{\text{beam}}^*$ corresponds to the characteristic propagation length of the EM field outside the pumped active medium and may be evaluated as $d_{\text{beam}}^* \sim v_{\text{group}}/\gamma_a$.
 For the parameters of the DFB laser we have $d_{\text{beam}}^* \sim 10 \mu \text{m}$, which is in agreement with the full model, Fig.~\ref{fig2}b.
@@
 To summarize, we have studied the temporal dynamics of a two-dimensional plasmonic DFB laser under ultrashort pulse pumping.
-Relying on numerical simulations and analytical calculations, we have shown that the response time of such a laser for typical experimental parameters can be decreased to one picosecond, which corresponds to a modulation frequency of $1$ THz.
+Relying on numerical simulations and analytical calculations, we have shown that the response time of such a laser for experimental parameters from ~\cite{TennerJOpt,TennerACSPhot} can be decreased to one picosecond, which corresponds to a modulation frequency of $1$ THz.
 This value is close to the upper limit for plasmonic lasers (the total response time can not be shorter than the time of the third stage, which is determined by the losses in the cavity, and is $ \sim 1$ ps).
@@
 In the plasmonic cavity, the inequality $\gamma_a \gg \gamma_D$ is always hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds.
-We believe that further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.
+Further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.
 
-Thus, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
+To sum up, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
 This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
@@
 
-\section*{Appendix A}
-To normalize the electric field ${{\bf{E}}_j}({{\bf{r}}_m})$, we equate the energy of one quantum to the energy of the electric field in the resonator:
+\section*{Appendix}
+To find the electric field ${{\bf{E}}_j}({{\bf{r}}_m})$  'per one quantum', we equate the energy of one quantum to the energy of the EM field in the resonator:
 \begin{equation}
@@
 \end{equation}
-where $\varepsilon (\omega, \textbf{r})$  is the dielectric constant of the active plasmonic structure.
+where $\varepsilon (\omega, \textbf{r})$  is the dielectric constant.
 
@@
 \end{equation}
-where $A_{0j}$ is a real normalization constant for \textit{j}th mode; ${{\bf{e}}_j}\left( {x,y,z} \right) = {{\bf{e}}_j}\left( {x + {L_x},y + {L_y},z} \right)$  and ${{\bf{h}}_j}\left( {x,y,z} \right) = {{\bf{h}}_j}\left( {x + {L_x},y + {L_y},z} \right)$, where $L_x$ and $L_y$ are periods of the plasmonic structure along $x$-axis and $y$-axis, respectively.
+where $A_{0j}$ is a real normalization constant for \textit{j}th mode; ${{\bf{e}}_j}\left( {x,y,z} \right) = {{\bf{e}}_j}\left( {x + {L_x},y + {L_y},z} \right)$  and ${{\bf{h}}_j}\left( {x,y,z} \right) = {{\bf{h}}_j}\left( {x + {L_x},y + {L_y},z} \right)$, $L_x$ and $L_y$ are periods of the plasmonic structure along $x$-axis and $y$-axis, respectively.
 
-The functions ${{\bf{e}}_j}\left(\textbf{r}\right)$ and ${{\bf{h}}_j}\left(\textbf{r}\right)$ can be find within the couple-mode theory, see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat of \cite{nefedkin2018acsphot}.
+The functions ${{\bf{e}}_j}\left(\textbf{r}\right)$ and ${{\bf{h}}_j}\left(\textbf{r}\right)$ can be found within the couple-mode theory, see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat of \cite{nefedkin2018acsphot}.
 In this approach, the eigenmodes of the plasmonic structure with holes are expressed as linear combinations of the eigenmodes of the plasmonic structure without holes \cite{TennerJOpt,TennerACSPhot}.
@@
 The in-plane electric field, $e_{\parallel}$, is in the direction of the mode's propagation. 
-The out-of-plane component of the electric field, $e_{\perp}$, is much greater than the in-plane component, $e_{\parallel}$ \cite{TennerJOpt,TennerACSPhot} and so it plays main role at the interaction of the modes with the active medium.
+The out-of-plane component of the electric field, $e_{\perp}$, is much greater than the in-plane component, $e_{\parallel}$ \cite{TennerJOpt,TennerACSPhot} and so it plays the main role in the interaction of the modes with the active medium.
 
-The eigenmodes of the plasmonic structure with the holes we find within the coupled-mode theory with the experimental values of the scattering amplitudes, see~\cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat.
+The eigenmodes of the plasmonic structure with the holes have been found within the coupled-mode theory with the experimental values of the scattering amplitudes, see~\cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat.
 of \cite{nefedkin2018acsphot}.
@@
 $\begin{array}{cccc}
-\text{The number of the dispersive curve} & e_{\parallel} & \
+\text{The number of the dispersion curve} & e_{\parallel} & \
 h_{\parallel} & \
@@
 For this reason, the modes that have a minimum value of $h_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $h_{\parallel}$ at the holes.
-Note that in the CW regime, the lasing takes place at the modes with small radiation losses and high quality factors lying on the 2nd dispersion curve, \cite{TennerJOpt,nefedkin2018acsphot}.
-\\
 
@@
 \end{equation}
-where $\Omega _{0jm}$ can be expressed as
+where $\Omega _{0jm}$ has the form
 \begin{equation}
@@
 {\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}}}{\left|\textbf{e}_{j}\right|_{max}^2}$ is mode volume of \textit{j}th mode.
+Thus, $\Omega_{0jm}^2 \sim 1/V_{jmode}$, and, as a consequence, is proportional to the Purcell factor of $j$th mode ~\cite{ZhouNatNano}.
 Also, ${f_j}\left( {{{\bf{r}}_m}} \right) = \left( {{{\bf{d}}_m} \cdot {{\bf{e}}_j}\left( {{{\bf{r}}_m}} \right)} \right)/(\left|\bf{d}_m\right| \left|\textbf{e}_{j}\right|_{max})$ is a periodic function with plasmonic lattice cell period.

### 2019-05-06 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
The second response

@@
 To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
-Although in our simulations we take into account several hundred of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response.
+In our simulations we take into account several hundred of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium.
+One can see that this distribution does not change in the first and second stages of the system response.
 In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field outside the pumped area and the stimulated amplification of the EM field inside the pumped area.
@@
 Furthermore, we assume that the pumped active medium interacts only with such a single collective mode.
-Within these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
+Within these approximations, the system dynamics can be described by the following system of equations \cite{SiegmanLasers}:
 \begin{gather} 
@@
 We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
+Note that the finite size of the pump beam can lead to appearing of edge effects and influence on the mode quality factor, see \cite{hakala2017lasing}.
+In the single-mode model, we do not take into account these effects.
 The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.

### 2019-05-06 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Second response, revised

@@
 We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
-Note that the finite size of the pump beam can lead to appearing of edge effects and influence on the mode quality factor, see \cite{hakala2017lasing}.
+Note that the finite size of the pump beam can lead to the appearance of edge effects and influence on the mode quality factor, see \cite{hakala2017lasing}.
 In the single-mode model, we do not take into account these effects.
@@
 
-To sum up, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
+In consequense, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
 This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.

### 2019-05-07 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Second revision, final version

@@
 $k_\text{pl}(\omega_{\sigma}) \approx G$, where $k_\text{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
-The linewidth of the active medium is about $100 \text{nm}$.
+The linewidth of the active medium is about $100$ $\text{nm}$.
 Similar plasmonic DFB lasers have been implemented in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and their parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
@@
 To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
-In our simulations we take into account several hundred of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium.
-One can see that this distribution does not change in the first and second stages of the system response.
+In Eqs.
+(\ref{eq1})-(\ref{eq3}), we take into account several hundred of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium.
+Numerical simulations show that this distribution does not change in the first and second stages of the system response.
 In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field outside the pumped area and the stimulated amplification of the EM field inside the pumped area.
@@
 Note that the finite size of the pump beam can lead to the appearance of edge effects and influence on the mode quality factor, see \cite{hakala2017lasing}.
-In the single-mode model, we do not take into account these effects.
+In the single-mode model, we do not take into account these effects for simplicity.
+We will further see that this simple model is in good agreement with the numerical simulations of Eqs. (\ref{eq1})-(\ref{eq3}).
 The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.
@@
 \begin{equation} 
-{t_{delay}} = (2\gamma _a)^{ - 1}{\left( {\xi  - 1} \right)^{ - 1}}\ln \left( {{N_{\text{at}}}{{\left( {\xi - 1} \right)}^2}/{\xi ^2}} \right). \label{tdele}
+{t_\text{delay}} = (2\gamma _a)^{ - 1}{\left( {\xi  - 1} \right)^{ - 1}}\ln \left( {{N_{\text{at}}}{{\left( {\xi - 1} \right)}^2}/{\xi ^2}} \right). \label{tdele}
 \end{equation}
 At this time, a sharp decrease of $D$ appears, see Fig.~\ref{fig4}b.
-Note that the assumption that the evolution of $D$ is much slower than the evolution of $n$ is valid when $t_{delay} \gg \gamma_a^{-1} \xi^{-1}$, i.e., $\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/{\xi ^2}} \right) \gg 1$.
+Note that the assumption that the evolution of $D$ is much slower than the evolution of $n$ is valid when $t_\text{delay} \gg \gamma_a^{-1} \xi^{-1}$, i.e., $\ln \left( {{N_{\text{at}}}{{\left( {\xi  - 1} \right)}^2}/{\xi ^2}} \right) \gg 1$.
 When $\xi \gg 1$, the assumption is correct if $N_{\text{at}} \gg 1$, which is true for the considered system.
@@
 
-In consequense, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
+To conclude we have shown that there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time.
 This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m.
